### 2023-01-22 manuscriptNature.Rmd
Minor edits

---
title: "Different rules for binocular combination of luminance flicker in cortical and subcortical pathways"
author: "Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker"
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  pdf_document:
    toc: no
  word_document: default
  html_document: default
bibliography: references.bib
csl: nature-neuroscience.csl
---

```{r setup, include=FALSE}

processdata <- 0  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data
# 2 - average data across participants, do bootstrapping and modelling
# 3 - download all raw data and analyse (requires 48GB of storage)

# check which packages are installed, install the missing ones, and activate
packagelist <- c('knitr','remotes','tictoc','R.matlab','bookdown','grImport','png','tiff','pals','ez','gtools','signal','boot','quickpsy','rstan','coda','parallel','osfr') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
if (!'FourierStats' %in% installed.packages()[,1]){remotes::install_github("bakerdh/FourierStats")}
packagelist <- c(packagelist,'FourierStats')
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

knitr::opts_chunk$set(echo = TRUE)

# helper function to make colours transparent
addalpha <- function(col, alpha=1){apply(sapply(col, col2rgb)/255, 2, function(x) rgb(x[1], x[2], x[3], alpha=alpha))}

v4Interp <- function(df, xo, yo, rmax = .75, gridRes = 67) {
  ## Create a function to perform Matlab's v4 interpolation.
  ## Takes as input a data-frame with columns x, y, and z (x co-ordinates, y co-ordinates, and amplitude)
  ## and variables xo and yo, the co-ordinates which will be use to create a grid for interpolation
  xo <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
  yo <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
  xy <- df$x + df$y*sqrt(as.complex(-1))
  d <- matrix(rep(xy,length(xy)),nrow = length(xy), ncol = length(xy))
  d <- abs(d - t(d))
  diag(d) <- 1
  g <- (d^2) * (log(d)-1)   # Green's function.
  diag(g) <- 0
  weights <- qr.solve(g,df$z)
  xy <- t(xy)
  outmat <- matrix(nrow = gridRes,ncol = gridRes)
  for (i in 1:gridRes){
    for (j in 1:gridRes) {
      test4 <- abs((xo[i,j] + sqrt(as.complex(-1))*yo[i,j]) - xy)
      g <- (test4^2) * (log(test4)-1)
      outmat[i,j] <- g %*% weights}}
  outDf <- data.frame(x = xo[,1],outmat)
  names(outDf)[1:length(yo[1,])+1] <- yo[1,]
  return(outDf)}

anovacirc2way <- function(data, groupA=NULL, groupB=NULL, participant=NULL){

  grouplabelsA <- groupA
  grouplabelsB <- groupB
  participantlabels <- participant
  datavals <- data

  grouplabelsA <- as.factor(grouplabelsA)
  factorlistA <- levels(grouplabelsA)
  grouplabelsB <- as.factor(grouplabelsB)
  factorlistB <- levels(grouplabelsB)
  participantlabels <- as.factor(participantlabels)
  participantlist <- levels(participantlabels)

  grandmean <- mean(datavals)
  grandvar <- sum(abs(datavals - grandmean)^2)/(length(datavals)-1)

    SST <- grandvar*(length(datavals) - 1)

    SSW <- 0
    for (n in 1:nlevels(participantlabels)){SSW <- SSW + sum(abs(datavals[which(participantlabels==participantlist[n])] - mean(datavals[which(participantlabels==participantlist[n])]))^2)}
    dfW <- 2*(nlevels(participantlabels)*(nlevels(grouplabelsA)*nlevels(grouplabelsB)-1))

    SSBet <- SST - SSW  # between subjects SS - not actually used
    SSM <- 0
    for (a in 1:nlevels(grouplabelsA)){
      for (b in 1:nlevels(grouplabelsB)){
        i <- as.numeric(((grouplabelsA==grouplabelsA[a])))
        j <- as.numeric(((grouplabelsB==grouplabelsB[b])))
        indices <- which((i*j)>0)
        SSM <- SSM + (nlevels(participantlabels)*sum(abs(datavals[indices] - grandmean)^2))
      }
    }
    dfM <- 2*((nlevels(grouplabelsA)*nlevels(grouplabelsB))-1)

    SSR <- 0
    for (a in 1:nlevels(grouplabelsA)){
      for (b in 1:nlevels(grouplabelsB)){
        i <- as.numeric(((grouplabelsA==grouplabelsA[a])))
        j <- as.numeric(((grouplabelsB==grouplabelsB[b])))
        indices <- which((i*j)>0)
        condvar <- sum(abs(datavals[indices] - mean(datavals[indices]))^2)/(length(indices)-1)
        SSR <- SSR + (nlevels(participantlabels)-1)*condvar
      }
    }    
    dfR <- 2*(nlevels(participantlabels)-1)*(nlevels(grouplabelsA)*nlevels(grouplabelsB))

    groupmeansA <- NULL
    for (n in 1:nlevels(grouplabelsA)){groupmeansA[n] <- mean(datavals[which(grouplabelsA==factorlistA[n])])}
    SSA <- 0
    for (n in 1:nlevels(grouplabelsA)){SSA <- SSA + length(which(grouplabelsA==factorlistA[n]))*abs(groupmeansA[n]-grandmean)^2}
    dfA <- 2*(nlevels(grouplabelsA)-1)

    groupmeansB <- NULL
    for (n in 1:nlevels(grouplabelsB)){groupmeansB[n] <- mean(datavals[which(grouplabelsB==factorlistB[n])])}
    SSB <- 0
    for (n in 1:nlevels(grouplabelsB)){SSB <- SSB + length(which(grouplabelsB==factorlistB[n]))*abs(groupmeansB[n]-grandmean)^2}
    dfB <- 2*(nlevels(grouplabelsB)-1)

    SSAB <- SSW - (SSA + SSB)
    dfAB <- 2 * (nlevels(grouplabelsA)-1) * (nlevels(grouplabelsB)-1)

    MSA <- SSA/dfA
    MSB <- SSB/dfB
    MSAB <- SSAB/dfAB
    MSR <- SSR/dfR

    FratioA <- MSA/MSR
    pvalA <- pf(FratioA,df1=dfA,df2=dfR,lower.tail=FALSE)
    FratioB <- MSB/MSR
    pvalB <- pf(FratioB,df1=dfB,df2=dfR,lower.tail=FALSE)
    FratioAB <- MSAB/MSR
    pvalAB <- pf(FratioAB,df1=dfAB,df2=dfR,lower.tail=FALSE)

    method <- 'Factorial repeated measures ANOVA^2-circ'
    output <- data.frame(FratioA,pvalA,FratioB,pvalB,FratioAB,pvalAB,dfA,dfB,dfAB,dfR,method)

  return(output)
}

colpal <- c('#FE5000','#8783CF','#228B22','#808080','#6d008b','#8B8000')
targetelectrodes <- c('Oz','POz','O1','O2')

localdir <- 'local/'    # all files are stored in the project directory /local/ which git is told to ignore
if (!file.exists(localdir)){dir.create(localdir)}   # create a local directory to store data and outputs
rawdir <- 'local/rawdata/'
if (!file.exists(rawdir)){dir.create(rawdir)}   # create a local directory to store raw data
figdir <- 'Figures/'
if (!file.exists(figdir)){dir.create(figdir)}   # create a local directory to store figures
datadir <- 'local/processeddata/'
if (!file.exists(datadir)){dir.create(datadir)}   # create a local directory to store processed data

# cut off for data inclusion (Mahalanobis distance for complex data)
SDthresh <- 3
nbootstraps <- 1000

```

# Abstract

How does the human brain combine information across the eyes? It has been known for many years that cortical normalisation mechanisms implement ‘ocularity invariance’: equalising neural responses to spatial patterns presented either monocularly or binocularly. Here we used a novel combination of electrophysiology, psychophysics, pupillometry and computational modelling to ask whether this invariance also holds for flickering luminance stimuli with no spatial contrast. We find dramatic violations of ocularity invariance for these stimuli, both in cortex and also in the subcortical pathways that govern pupil diameter. Specifically, we find substantial binocular facilitation in both pathways with the effect being strongest in cortex. Near-linear binocular additivity (instead of ocularity invariance) was also found using a perceptual luminance matching task. Ocularity invariance is therefore not a ubiquitous feature of visual processing, and the brain appears to repurpose a generic normalisation algorithm for different visual functions by adjusting the amount of interocular suppression.

# Introduction

The brain must combine information across multiple sensory inputs to derive a coherent percept of the external world. This involves a process of signal combination both within [@Baker2017] and between [@Ernst2002] the senses. Binocular vision is a useful test case for signal combination, as the inputs to the two eyes overlap substantially (in species with forward-facing eyes), and the neural locus is well-established [@Hubel1962]. Much of our knowledge about binocular combination derives from studies on the contrast response of the 'canonical' visual pathway, in which signals pass from the eyes to primary visual cortex (V1), via the lateral geniculate nucleus (LGN) [@Purves2008]. However, signals are also combined across the eyes in the network of subcortical nuclei that govern pupil diameter in response to absolute light levels [@McDougal2008] and much less is known about the computations that operate in these subcortical pathways. Our primary purpose here is to investigate the computations governing signal combination in these two anatomically distinct pathways in response to luminance changes.

For pattern vision, binocular presentation confers greater sensitivity to low contrast targets than monocular presentation. This is known as binocular summation[@Campbell1965; @Baker2018], with summation ratios (the relative improvement under binocular presentation) at detection threshold lying between $\sqrt{2}$ and 2. This advantage is lost at high stimulus intensities, where both psychophysical performance (contrast discrimination thresholds) [@Legge1984; @Meese2006] and neural activity [@Moradi2009; @Baker2017] are approximately equal for monocular and binocular presentation. Contemporary models of binocular vision [@Ding2006; @Meese2006] advocate a process of interocular suppression that normalizes the two eyes inputs at high contrasts and negates the binocular advantage. This is consistent with our everyday experience of 'ocularity invariance' [@Baker2007]: perceived contrast does not change when one eye is opened and closed.

The pupillary light reflex is an automatic constriction of the iris sphincter muscles in response to increases in light levels, which causes the pupil to shrink [@McDougal2008]. There is a clear binocular component to this reflex, as stimulation of one eye still causes constriction of the other eye's pupil (termed the consensual response; @Wyatt1981). Importantly, the neuroanatomical pathway involved completely bypasses the canonical cortical pathway (retina to V1), instead involving a network of subcortical nuclei, including the Pretectal Olivary nucleus, Superior Cervical ganglion, and Edinger-Westphal nucleus [@Angee2021; @Mathot2018; @McDougal2008; @Wang2015]. To account for the consensual response, these brain regions must  combine information from the left and right eyes [@Doesschate1967], yet the computation that achieves this is unclear.

To address this shortcoming, we designed an experiment that allowed us to simultaneously record electrophysiological and pupillometric responses to monocular and binocular stimuli. This novel paradigm allowed us to probe both cortical (using EEG) and subcortical (using a binocular eyetracker) pathways simultaneously in response to flickering light, and make quantitative comparisons between them. Periodic flicker entrains both neural [@Norcia2015] and pupil [@Spitschan2014] responses at the flicker frequency, enabling precise estimation of response amplitudes in the Fourier domain. We followed up our main experiment with additional exploration of the effect of stimulus frequency, and a psychophysical matching experiment measuring perceived flicker intensity (i.e. temporal contrast). The results are interpreted using a hierarchical Bayesian computational model of binocular vision, and reveal that subcortical pathways implement stronger interocular suppression than the canonical cortical pathway.

# Results

## Experiment 1

```{r include=FALSE, results='hide'}

# (download and) analyse individual participant data for experiment 1

if (processdata > 2){

legaltriggers <- 1

timeseq <- seq(1/120,10,length.out=120*10)
timeseq2 <- seq(1/120,14,length.out=120*14)
EEGtimes <- seq(1/1000,10,1/1000)
targetindex <- (2*10)+1
maskindex <- (1.6*10)+1
targetindex2 <- (2*2*10)+1
maskindex2 <- (2*1.6*10)+1

  for (participant in 1:30){

    # only download and process data if this has not been done already for this participant
    if (!file.exists(paste0(datadir,'P',100+participant,'summary.RData'))){

      # if there is no directory for this participant's data
      if (!file.exists(paste0(rawdir,'P',100+participant))){
        dir.create(paste0(rawdir,'P',100+participant))

       d <- dir(paste0(rawdir,'P',100+participant))

       if (length(d)==0){

         # if the tar file of this participant's data doesn't exist
         if (!file.exists(paste0(rawdir,'P',100+participant,'.tar'))){
           # download it from OSF
           if (!exists('osffiles')){
           osfnode <- 'x8u4v'
           osfproject <- osf_retrieve_node(osfnode)
          osffiles <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffiles$name==paste0('P',100+participant,'.tar'))
          osf_download(osffiles[fid,],rawdir,progress=TRUE)
         }

         # then unzip the tar file
         untar(paste0(rawdir,'P',100+participant,'.tar'),exdir=paste0(rawdir,'P',100+participant))

         # and delete the tar file to save storage space
         file.remove(paste0(rawdir,'P',100+participant,'.tar'))
       }

      d <- dir(paste0(rawdir,'P',100+participant),full.names=TRUE)

      pupiltargets <- array(0,dim=c(3,2,60))
pupilmasks <- pupiltargets
pupilwaveforms <- array(0,dim=c(3,2,60,120*14))
pupilspectra <- array(0,dim=c(3,2,60,300))
EEGtargets <- array(0,dim=c(3,64,60))
EEGmasks <- EEGtargets
EEGtargets2 <- EEGtargets
EEGmasks2 <- EEGtargets
EEGwaveforms <- array(0,dim=c(3,64,60,1000*14))
EEGspectra <- array(0,dim=c(3,64,60,300))

psychopyfiles <- dir(path=paste0(rawdir,'P',100+participant),pattern=paste0('P',100+participant,'_CRFstudy'), full.names = TRUE)
infofiles <- dir(path=paste0(rawdir,'P',100+participant),pattern='*_info.csv', full.names = TRUE)
EEGfiles <- dir(path=paste0(rawdir,'P',100+participant),pattern='*_EEG.csv.gz', full.names = TRUE)
pupilfiles <- dir(path=paste0(rawdir,'P',100+participant),pattern='*_pupil_positions.csv', full.names = TRUE)

for (block in 1:length(EEGfiles)){

    infofile <- read.csv(infofiles[block])
  pupilstarttime <- as.numeric(as.character(infofile[4,2]))
  localstarttime <- as.numeric(as.character(infofile[5,2]))

  psychoutput <- read.csv(psychopyfiles[block])

  trialtimes <- psychoutput$trialonset - pupilstarttime
  condorder <- psychoutput$condition

    EEGdata <- read.csv(EEGfiles[block])
  electrodes <- colnames(EEGdata)

      triggertimes <- NULL
    counter <- 0
    lasttrigger <- -10000
    for (n in 1:nrow(EEGdata)){
      if(EEGdata$Trigger[n] %in% legaltriggers){
        if(n>(lasttrigger+10000)){
        counter <- counter + 1
        triggertimes[counter] <- n
        lasttrigger <- n
      }}
    }

    for (cond in 1:60){
      for (ch in 1:64){
        trial <- EEGdata[(triggertimes[cond]+(2001:12000)),ch+2]
        fspec <- (fft(trial)/length(trial))
        EEGtargets[block,ch,condorder[cond]] <- fspec[targetindex]
        EEGmasks[block,ch,condorder[cond]] <- fspec[maskindex]
        EEGtargets2[block,ch,condorder[cond]] <- fspec[targetindex2]
        EEGmasks2[block,ch,condorder[cond]] <- fspec[maskindex2]
        EEGspectra[block,ch,condorder[cond],1:300] <- fspec[1:300]

        trial <- EEGdata[(triggertimes[cond]+(-999:13000)),ch+2]
        EEGwaveforms[block,ch,condorder[cond],] <- trial - mean(trial[1:1000])
      }}

  pdata <- read.csv(pupilfiles[block])
  pdata2 <- pdata[,c(1,3,4,14)]
  pdata2[,1] <- pdata2[,1] - localstarttime
  pdata2 <- pdata2[which(pdata2[,3]>0.0),]

  for (eye in 1:2){
    eyedata <- pdata2[which(pdata2[,2]==(2-eye)),]
    for (cond in 1:60){
      a <- which(eyedata[,1]>trialtimes[cond]+2)
      b <- which(eyedata[,1]<(trialtimes[cond]+12))
      i <- intersect(a,b)
      trial <- eyedata[i,]
      trial[,1] <- trial[,1] - trial[1,1]
      if (nrow(trial)>3){
        resampled <- interp1(trial[,1],trial[,4],timeseq,method='linear',extrap=TRUE)
        fspec <- (fft(resampled)/length(resampled))
        pupiltargets[block,eye,condorder[cond]] <- fspec[targetindex]
        pupilmasks[block,eye,condorder[cond]] <- fspec[maskindex]
        pupilspectra[block,eye,condorder[cond],1:300] <- fspec[1:300]
      }

      a <- which(eyedata[,1]>(trialtimes[cond]-1))
      b <- which(eyedata[,1]<(trialtimes[cond]+13))
      i <- intersect(a,b)
      trial <- eyedata[i,]
      trial[,1] <- trial[,1] - trial[1,1]
      if (nrow(trial)>3){
        resampled <- interp1(trial[,1],trial[,4],timeseq2,method='linear',extrap=TRUE)
        pupilwaveforms[block,eye,condorder[cond],] <- resampled - mean(resampled[1:120])
      }

      }}

}

cleanmeansP <- matrix(0,nrow=6,ncol=5)
cleanmasksP <- matrix(0,nrow=6,ncol=5)
cleanmeansE <- array(0,dim=c(64,6,5))
cleanmasksE <- array(0,dim=c(64,6,5))
cleanmeansE2 <- array(0,dim=c(64,6,5))
cleanmasksE2 <- array(0,dim=c(64,6,5))

for (cond in 1:6){
  startindex <- (10*(cond-1))
  for (level in 1:5){
    temp <- c(pupiltargets[,,startindex+level],pupiltargets[,,startindex+level+5])
        cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmeansP[cond,level] <- mean(temp[i])}

    temp <- c(pupilmasks[,,startindex+level],pupiltargets[,,startindex+level+5])
        cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmasksP[cond,level] <- mean(temp[i])}

    for (ch in 1:64){
    temp <- c(EEGtargets[,ch,startindex+level],EEGtargets[,ch,startindex+level+5])
            cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmeansE[ch,cond,level] <- mean(temp[i])}

    temp <- c(EEGmasks[,ch,startindex+level],EEGmasks[,ch,startindex+level+5])
                cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmasksE[ch,cond,level] <- mean(temp[i])}

    temp <- c(EEGtargets2[,ch,startindex+level],EEGtargets2[,ch,startindex+level+5])
                cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmeansE2[ch,cond,level] <- mean(temp[i])}

    temp <- c(EEGmasks2[,ch,startindex+level],EEGmasks2[,ch,startindex+level+5])
                cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    cleanmasksE2[ch,cond,level] <- mean(temp[i])}   
    }

  }
}

electrodeindices <- match(targetelectrodes,electrodes)-2

meanwavesP <- apply(pupilwaveforms[,,c(15,20),],4,mean,na.rm=TRUE)
meanwavesE <- apply(EEGwaveforms[,electrodeindices,c(15,20),],4,mean,na.rm=TRUE)
meanspectraP <- apply(pupilspectra[,,c(15,20),],4,mean,na.rm=TRUE)
meanspectraE <- apply(EEGspectra[,electrodeindices,c(15,20),],4,mean,na.rm=TRUE)
save(file=paste(datadir,'P',100+participant,'summary.RData',sep=''),list=c('cleanmeansP','cleanmeansE','cleanmeansE2','cleanmasksP','cleanmasksE','cleanmasksE2','meanspectraP','meanspectraE','meanwavesP','meanwavesE','electrodes'))

      }

      }

    }
  }

```

```{r include=FALSE, results='hide'}

# load in individual participant data and average, save group data

if (processdata > 1){

         if (!file.exists(paste0(datadir,'headerfile.csv'))){
           # download it from OSF
           if (!exists('osffiles')){
           osfnode <- 'x8u4v'
           osfproject <- osf_retrieve_node(osfnode)
          osffiles <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffiles$name=='headerfile.csv')
          osf_download(osffiles[fid,],datadir,progress=TRUE)
          }

  hdata <- read.csv(paste0(datadir,'headerfile.csv'))

for (participant in 1:30){
         # if the summary file of this participant's data doesn't exist
         if (!file.exists(paste0(datadir,'P',100+participant,'summary.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name==paste0('P',100+participant,'summary.RData'))
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }
}

d <- dir(datadir,pattern='*summary.RData',full.names=TRUE)

allmasksP <- array(0,dim=c(length(d),6,5))
allmeansP <- allmasksP
allmasksE <- array(0,dim=c(length(d),64,6,5))
allmeansE <- allmasksE
allmeansE2 <- allmeansE
allmasksE2 <- allmasksE
allspectraP <- array(0,dim=c(length(d),300))
allspectraE <- array(0,dim=c(length(d),300))
allwavesP <- array(0,dim=c(length(d),1680))
allwavesE <- array(0,dim=c(length(d),14000))

  for (s in 1:length(d)){

    load(d[s])

  allmasksP[s,,] <- cleanmasksP
  allmeansP[s,,] <- cleanmeansP
  allspectraP[s,] <- meanspectraP
  allspectraE[s,] <- meanspectraE
  allwavesP[s,] <- meanwavesP
  allwavesE[s,] <- meanwavesE
  allmeansE[s,,,] <- cleanmeansE
  allmasksE[s,,,] <- cleanmasksE
  allmeansE2[s,,,] <- cleanmeansE2
  allmasksE2[s,,,] <- cleanmasksE2

  }

electrodeindices <- match(targetelectrodes,electrodes)-2

allspectraP <- allspectraP[-c(2,4,13,20,26),]

cleanmeansP <- matrix(0,nrow=6,ncol=5)
cleanmeansPCI <- array(0,c(2,6,5))
cleanmasksP <- matrix(0,nrow=6,ncol=5)
cleanmasksPCI <- array(0,c(2,6,5))
cleanmeansEav <- matrix(0,nrow=6,ncol=5)
cleanmeansEavCI <- array(0,c(2,6,5))
cleanmasksEav <- matrix(0,nrow=6,ncol=5)
cleanmasksEavCI <- array(0,c(2,6,5))
cleanmeansE2av <- matrix(0,nrow=6,ncol=5)
cleanmeansE2avCI <- array(0,c(2,6,5))
cleanmasksE2av <- matrix(0,nrow=6,ncol=5)
cleanmasksE2avCI <- array(0,c(2,6,5))
cleanmeansE <- array(0,dim=c(64,6,5))
cleanmeansECI <- array(0,c(2,64,6,5))
cleanmasksE <- array(0,dim=c(64,6,5))
cleanmasksECI <- array(0,c(2,64,6,5))
cleanmeansE2 <- array(0,dim=c(64,6,5))
cleanmeansE2CI <- array(0,c(2,64,6,5))
cleanmasksE2 <- array(0,dim=c(64,6,5))
cleanmasksE2CI <- array(0,c(2,64,6,5))

for (cond in 1:6){
  for (level in 1:5){

    temp <- allmeansP[,cond,level]
    cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmeansP[cond,level] <- mbs$meanamp
    cleanmeansPCI[1,cond,level] <- mbs$lowerCI
    cleanmeansPCI[2,cond,level] <- mbs$upperCI
    }

    temp <- allmasksP[,cond,level]
    cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
          mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmasksP[cond,level] <- mbs$meanamp
    cleanmasksPCI[1,cond,level] <- mbs$lowerCI
    cleanmasksPCI[2,cond,level] <- mbs$upperCI
}

      temp <- apply(allmeansE[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmeansEav[cond,level] <- mbs$meanamp
    cleanmeansEavCI[1,cond,level] <- mbs$lowerCI
    cleanmeansEavCI[2,cond,level] <- mbs$upperCI
      }

temp <- apply(allmasksE[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmasksEav[cond,level] <- mbs$meanamp
    cleanmasksEavCI[1,cond,level] <- mbs$lowerCI
    cleanmasksEavCI[2,cond,level] <- mbs$upperCI
        }

temp <- apply(allmeansE2[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmeansE2av[cond,level] <- mbs$meanamp
    cleanmeansE2avCI[1,cond,level] <- mbs$lowerCI
    cleanmeansE2avCI[2,cond,level] <- mbs$upperCI
        }

      temp <- apply(allmasksE2[,electrodeindices,cond,level],1,mean)
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmasksE2av[cond,level] <- mbs$meanamp
    cleanmasksE2avCI[1,cond,level] <- mbs$lowerCI
    cleanmasksE2avCI[2,cond,level] <- mbs$upperCI
        }

    for (ch in 1:64){

      temp <- allmeansE[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmeansE[ch,cond,level] <- mbs$meanamp
    cleanmeansECI[1,ch,cond,level] <- mbs$lowerCI
    cleanmeansECI[2,ch,cond,level] <- mbs$upperCI
}

      temp <- allmasksE[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmasksE[ch,cond,level] <- mbs$meanamp
    cleanmasksECI[1,ch,cond,level] <- mbs$lowerCI
    cleanmasksECI[2,ch,cond,level] <- mbs$upperCI
        }

      temp <- allmeansE2[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmeansE2[ch,cond,level] <- mbs$meanamp
    cleanmeansE2CI[1,ch,cond,level] <- mbs$lowerCI
    cleanmeansE2CI[2,ch,cond,level] <- mbs$upperCI
        }

      temp <- allmasksE2[,ch,cond,level]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmasksE2[ch,cond,level] <- mbs$meanamp
    cleanmasksE2CI[1,ch,cond,level] <- mbs$lowerCI
    cleanmasksE2CI[2,ch,cond,level] <- mbs$upperCI
        }

    }}}

meanspectraP <- (1:300)*0
meanspectraE <- (1:300)*0
meanspectraPCI <- array(0,c(2,300))
meanspectraECI <- array(0,c(2,300))

for (f in 2:300){

        temp <- allspectraP[,f]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
meanspectraP[f] <- mbs$meanamp
meanspectraPCI[1,f] <- mbs$lowerCI
meanspectraPCI[2,f] <- mbs$upperCI
      }

        temp <- allspectraE[,f]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
meanspectraE[f] <- mbs$meanamp
meanspectraECI[1,f] <- mbs$lowerCI
meanspectraECI[2,f] <- mbs$upperCI
      }
}

meanwavesP <- NULL
meanwavesE <- NULL
meanwavesPCI <- array(0,dim=c(2,dim(allwavesP)[2]))
meanwavesECI <- array(0,dim=c(2,dim(allwavesE)[2]))
nsubjs <- dim(allwavesP)[1]
for (t in 1:dim(allwavesP)[2]){
  temp <- allwavesP[,t]
  i <- which(abs(temp-mean(temp))<(SDthresh*sd(temp)))
  meanwavesP[t] <- mean(temp[i])
  bspop <- NULL
  for (s in 1:nbootstraps){bspop[s] <- mean(sample(temp[i],length(i),replace=TRUE))}
  meanwavesPCI[,t] <- quantile(bspop,c(0.16,0.84),na.rm=TRUE)
}

for (t in 1:dim(allwavesE)[2]){
    temp <- allwavesE[,t]
  i <- which(abs(temp-mean(temp))<(SDthresh*sd(temp)))
  meanwavesE[t] <- mean(temp[i])
  bspop <- NULL
  for (s in 1:nbootstraps){bspop[s] <- mean(sample(temp[i],length(i),replace=TRUE))}
  meanwavesECI[,t] <- quantile(bspop,c(0.16,0.84),na.rm=TRUE)
}

save(file=paste0(datadir,'Averagedata.RData'),list=c('cleanmeansP','cleanmeansPCI','cleanmasksP','cleanmasksPCI','cleanmeansEav','cleanmeansEavCI','cleanmasksEav','cleanmasksEavCI','cleanmeansE2av','cleanmeansE2avCI','cleanmasksE2av','cleanmasksE2avCI','cleanmeansE','cleanmeansECI','cleanmasksECI','cleanmeansE2','cleanmeansE2CI','cleanmasksE2','cleanmasksE2CI','meanspectraP','meanspectraPCI','meanspectraE','meanspectraECI','meanwavesP','meanwavesE','meanwavesPCI','meanwavesECI','electrodes','allmeansP','allmeansE','allmeansE2'))

}

```

```{r include=FALSE, results='hide'}

# generate plot showing the group pupillometry data

         if (!file.exists(paste0(datadir,'Averagedata.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='Averagedata.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }

load(file=paste0(datadir,'Averagedata.RData'))

if (processdata > 0){

contrastsdB <- 20*log10(c(6,12,24,48,96))

plotlims <- c(15,40,0,0.03) 
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklocsy <- seq(0,0.03,0.01)    # locations of tick marks on y axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("CRF1p.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 2Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansP[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansP[cond,]),contrastsdB,cleanmeansPCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP[cond,]),contrastsdB,cleanmeansPCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
legend(22,0.03,c('Monocular','Binocular','Dichoptic'),pch=21:23,pt.bg=colpal[1:3],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.028,'(c)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF2p.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude at 2Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

condlist <- c(1,5,6)
pointlist <- c(21,24,25)
for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansP[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)  
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansP[condlist[cond],]),contrastsdB,cleanmeansPCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)  
  points(contrastsdB,abs(cleanmeansP[condlist[cond],]), pch = pointlist[cond], col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(22,0.03,c('Monocular','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.028,'(d)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF3p.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 1.6Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

for (cond in 4:6){
  lines(contrastsdB,abs(cleanmasksP[cond,]), col=colpal[cond], lwd=3, cex=0.5)   
  arrows(contrastsdB,abs(cleanmasksP[cond,]),contrastsdB,cleanmasksPCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksP[cond,]),contrastsdB,cleanmasksPCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)    
  points(contrastsdB,abs(cleanmasksP[cond,]), pch = pointlist[cond-3], col='black', bg=colpal[cond], cex=1.6, lwd=3) 
}
legend(22,0.03,c('Monocular 1.6Hz','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[4:6],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.028,'(e)',adj=0.5,cex=2.5)

dev.off()

postscript("SpecP.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 8)

frequencies <- (0:299)/10

plotlims <- c(1,6,0,0.03) 
ticklocsx <- seq(1,6,1)    # locations of tick marks on x axis
ticklocsy <- seq(0,0.03,0.01)    # locations of tick marks on y axis
ticklabelsx <-ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

polygon(frequencies[c(11:61,61:11)],c(meanspectraPCI[1,11:61],meanspectraPCI[2,61:11]),col=colpal[2],border=NA)

lines(frequencies[11:61],abs(meanspectraP[11:61]), col=colpal[2], lwd=3, cex=0.5)

text(1.1,0.028,'(b)',adj=0.5,cex=2.5)

dev.off()

postscript("timecourseP.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 8)

times <- seq((1/120)-1,13,1/120)
sinewave <- sin(2*times * 2*pi)
sinewave[1:120] <- 0
sinewave[1560:1680] <- 0

plotlims <- c(-1,13,-1,0.5) 
ticklocsx <- seq(-1,13,1)    # locations of tick marks on x axis
ticklocsy <- seq(-1,0.5,0.5)    # locations of tick marks on y axis
ticklabelsx <- ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

filt <- butter(3,0.083)
filtwaveP <- signal::filter(filt, meanwavesP)
filtU <- signal::filter(filt,meanwavesPCI[1,])
filtL <- signal::filter(filt,meanwavesPCI[2,])
polygon(times[c(1:length(times),length(times):1)],c(filtL,filtU[length(filtU):1]),col=colpal[2],border=NA)

lines(times, filtwaveP, col=colpal[2], lwd=1.5)

lines(times, (sinewave/10)-0.9, col='black', lwd=3)

text(-0.7,0.4,'(a)',adj=0.5,cex=2.5)

dev.off()

         if (!file.exists(paste0(localdir,'StimulusExample.png'))){
           # download it from OSF
           if (!exists('osffiles')){
           osfnode <- 'tbema'
           osfproject <- osf_retrieve_node(osfnode)
          osffiles <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffiles$name=='StimulusExample.png')
          osf_download(osffiles[fid,],localdir,progress=TRUE)
         }

stim <- readPNG('local/StimulusExample.png')

  PostScriptTrace('CRF1p.ps')
  p1 <- readPicture('CRF1p.ps.xml')
  PostScriptTrace('CRF2p.ps')
  p2 <- readPicture('CRF2p.ps.xml')
  PostScriptTrace('CRF3p.ps')
  p3 <- readPicture('CRF3p.ps.xml')
  PostScriptTrace('SpecP.ps')
  p4 <- readPicture('SpecP.ps.xml')
  PostScriptTrace('timecourseP.ps')
  p5 <- readPicture('timecourseP.ps.xml')

    for (n in 1:length(p4@paths)){
    temp <- class(p4@paths[n]$path)[1]
    if (pmatch(temp,"PictureFill",nomatch=0)){
      if (sum(col2rgb(p4@paths[n]$path@rgb))<765){p4@paths[n]$path@rgb <- addalpha(p4@paths[n]$path@rgb,alpha=0.2)}}}

      for (n in 1:length(p5@paths)){
    temp <- class(p5@paths[n]$path)[1]
    if (pmatch(temp,"PictureFill",nomatch=0)){
      if (sum(col2rgb(p5@paths[n]$path@rgb))<765){p5@paths[n]$path@rgb <- addalpha(p5@paths[n]$path@rgb,alpha=0.2)}}}

  pdf(paste0(figdir,"pupildata.pdf"), bg="transparent", height = 10, width = 15)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  aspratio <- 10/15  # this is the aspect ratio of the output pdf
  imwidth <- 0.2
  xstart <- 0.85
  ystart <- 0.72
  rasterImage(stim,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)

  grid.picture(p1,x=0.17,y=0.25,width=0.32,height=1)
  grid.picture(p2,x=0.5,y=0.25,width=0.32,height=1)
  grid.picture(p3,x=0.83,y=0.25,width=0.32,height=1)
  grid.picture(p5,x=0.25,y=0.7,width=0.5,height=1)
  grid.picture(p4,x=0.75,y=0.7,width=0.5,height=1)

  text(0.5,0.98,'Pupillometry',cex=4)

  dev.off()

  file.remove(c('CRF1p.ps','CRF2p.ps','CRF3p.ps','SpecP.ps','timecourseP.ps'))
  file.remove(c('CRF1p.ps.xml','CRF2p.ps.xml','CRF3p.ps.xml','SpecP.ps.xml','timecourseP.ps.xml'))

}

```

```{r include=FALSE, results='hide'}

# do statistics on the pupillometry data

alldata <- NULL
groupID <- NULL
allP <- NULL
counter <- 0
for (cont in 1:dim(allmeansP)[3]){
  for (p in 1:dim(allmeansP)[1]){
    counter <- counter + 1
    alldata[counter] <- allmeansP[p,3,cont]
    groupID[counter] <- cont
    allP[counter] <- p
  }
}
dichA2c <- anovacirc.test(alldata,group=groupID,participant=allP)
# groupID <- as.factor(groupID)
# participantID <- as.factor(participantID)
# adata <- data.frame(dichdata,groupID,participantID)
# dichA2c <- ezANOVA(data = adata, dv = dichdata, within=groupID, wid=participantID)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    for (cond in 1:2){
    counter <- counter + 1
    alldata[counter] <- allmeansP[p,cond,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- cond
    allP[counter] <- p
    }
  }
}
monbinPstats <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# monbinPstats <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    counter <- counter + 1
    alldata[counter] <- allmeansP[p,1,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- 1
    allP[counter] <- p

    counter <- counter + 1
    alldata[counter] <- allmeansP[p,3,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- 2
    allP[counter] <- p
  }
}
mondichPstats <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# mondichPstats <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    counter <- counter + 1
    alldata[counter] <- allmeansP[p,1,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- 1
    allP[counter] <- p

    counter <- counter + 1
    alldata[counter] <- allmeansP[p,5,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- 2
    allP[counter] <- p
  }
}

monbinPstatsX <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# monbinPstatsX <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    counter <- counter + 1
    alldata[counter] <- allmeansP[p,1,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- 1
    allP[counter] <- p

    counter <- counter + 1
    alldata[counter] <- allmeansP[p,6,lev]
    allgroupA[counter] <- lev
    allgroupB[counter] <- 2
    allP[counter] <- p
  }
}

mondichPstatsX <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# mondichPstatsX <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

```

The pupillometry results for our first experiment are summarised in Fig. \@ref(fig:pupildata). The group average waveform for binocular presentation is shown in Fig. \@ref(fig:pupildata)a. There is a substantial pupil constriction at stimulus onset, followed by visible oscillations at the flicker frequency (2Hz, see waveform at foot). The average Fourier spectrum is displayed in Fig. \@ref(fig:pupildata)b, and shows a clear spike at 2 Hz, but no evidence of a second harmonic response at 4Hz. These results demonstrate that our paradigm can evoke measurable steady-state pupil responses at 2Hz.

```{r pupildata, fig.cap="Summary of pupillometry results for N=30 participants. Panel (a) shows a group average waveform for binocular presentation (low pass filtered at 5Hz), with the driving signal plotted at the foot. Panel (b) shows the average Fourier spectrum, with an inset image illustrating the stimulus appearance (upper right). Panels (c,d) show contrast response functions at 2Hz for different conditions. Panel (e) shows contrast response functions at 1.6Hz for three conditions. Shaded regions and error bars indicate bootstrapped standard errors.", fig.align="center", echo=FALSE}

knitr::include_graphics(paste0(figdir,'pupildata.pdf'))

```

Fig. \@ref(fig:pupildata)c shows contrast response functions in response to stimuli flickering only at 2Hz. Response amplitudes increased monotonically with target contrast, confirming that our paradigm is suitable for measuring contrast-dependent differences in response (to our knowledge this is the first time this has been demonstrated). The amplitude of the binocular condition (blue squares) is consistently greater than that of the monocular condition (red circles) across all target contrasts. A $2\times5$ repeated measures $ANOVA^2_{circ}$ [@Baker2021] comparing these conditions revealed a significant main effect of target contrast (F(`r monbinPstats$dfA`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioA,digits=2)`, $p$ < 0.001), a significant effect of condition (F(`r monbinPstats$dfB`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioB,digits=2)`, $p$ < 0.001), and a significant interaction (F(`r monbinPstats$dfAB`,`r monbinPstats$dfR`) = `r round(monbinPstats$FratioAB,digits=2)`, $p$ < 0.001). The dichoptic condition begins at a much higher amplitude, owing to binocular combination of the target and high (48%) contrast mask, and then increases slightly with increasing target contrast (main effect of target contrast: F(`r dichA2c$dfM`,`r dichA2c$dfR`) = `r round(dichA2c$Fratio,digits=2)`, $p$ < `r round(dichA2c$pval,digits=3)`).

In Fig. \@ref(fig:pupildata)d, we plot responses to monocular target stimuli flickering at 2Hz, when the other eye viewed stimuli flickering at 1.6Hz (the red monocular-only data are replotted from Fig. \@ref(fig:pupildata)c for comparison). When the 1.6Hz component had the same contrast as the target (the binocular cross condition, shown in purple) responses were facilitated slightly at low contrasts, and suppressed at the highest target contrasts (interaction between contrast and condition: F(`r monbinPstatsX$dfAB`,`r monbinPstatsX$dfR`) = `r round(monbinPstatsX$FratioAB,digits=2)`, $p$ < 0.001). When the 1.6Hz component had a fixed contrast of 48% (the dichoptic cross condition, shown in yellow), responses were suppressed slightly across the contrast range (interaction between contrast and condition: F(`r mondichPstatsX$dfAB`,`r mondichPstatsX$dfR`) = `r round(mondichPstatsX$FratioAB,digits=2)`, $p$ < 0.001).

Fig. \@ref(fig:pupildata)e shows responses at 1.6Hz, for the same conditions, as well as for a condition in which a monocular stimulus flickered at 1.6Hz (grey circles). Surprisingly there again appears to be a slight facilitation effect in the binocular cross condition, particularly at lower contrasts. The dichoptic cross condition does not show clear modulation with target contrast.

Fig. \@ref(fig:EEGdata) shows equivalent results, measured contemporaneously using EEG. Fig. \@ref(fig:EEGdata)a shows the group average waveform for binocular presentation, and Fig. \@ref(fig:EEGdata)b shows the Fourier spectrum for binocular presentation, both averaged across four posterior electrodes (_Oz_, _POz_, _O1_ and _O2_, marked on the inset scalp plots). Unlike for the pupillometry data, there are clear responses at both the first harmonic frequency (2Hz), and also the second harmonic frequency (4Hz). We therefore derived contrast response functions at both first and second harmonic frequencies.

```{r include=FALSE, results='hide'}

# generate plot showing the group EEG data

contrastsdB <- 20*log10(c(6,12,24,48,96))

         if (!file.exists(paste0(datadir,'headerfile.csv'))){
           # download it from OSF
           if (!exists('osffiles')){
           osfnode <- 'x8u4v'
           osfproject <- osf_retrieve_node(osfnode)
          osffiles <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffiles$name=='headerfile.csv')
          osf_download(osffiles[fid,],datadir,progress=TRUE)
          } 
hdata <- read.csv(paste0(datadir,'headerfile.csv'))

         if (!file.exists(paste0(datadir,'Averagedata.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='Averagedata.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }

load(file=paste0(datadir,'Averagedata.RData'))

electrodeindices <- match(targetelectrodes,electrodes)-2

if (processdata > 0){

plotlims <- c(15,40,0,0.8)  
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklocsy <- seq(0,0.8,0.2)    # locations of tick marks on y axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("CRF1e.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV) at 2Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansEav[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansEav[cond,]),contrastsdB,cleanmeansEavCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansEav[cond,]),contrastsdB,cleanmeansEavCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
legend(22,0.8,c('Monocular','Binocular','Dichoptic'),pch=21:23,pt.bg=colpal[1:3],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(c)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF2e.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude (µV) at 2Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

condlist <- c(1,5,6)
pointlist <- c(21,24,25)
for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansEav[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)  
  arrows(contrastsdB,abs(cleanmeansEav[condlist[cond],]),contrastsdB,cleanmeansEavCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansEav[condlist[cond],]),contrastsdB,cleanmeansEavCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansEav[condlist[cond],]), pch = pointlist[cond], col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(22,0.8,c('Monocular','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(d)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF3e.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude (µV) at 1.6Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 4:6){
  lines(contrastsdB,abs(cleanmasksEav[cond,]), col=colpal[cond], lwd=3, cex=0.5)   
  arrows(contrastsdB,abs(cleanmasksEav[cond,]),contrastsdB,cleanmasksEavCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksEav[cond,]),contrastsdB,cleanmasksEavCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmasksEav[cond,]), pch = pointlist[cond-3], col='black', bg=colpal[cond], cex=1.6, lwd=3) 
}
legend(22,0.8,c('Monocular 1.6Hz','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[4:6],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(e)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF1e2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV) at 4Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansE2av[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  arrows(contrastsdB,abs(cleanmeansE2av[cond,]),contrastsdB,cleanmeansE2avCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  if (cond!=2){
  arrows(contrastsdB,abs(cleanmeansE2av[cond,]),contrastsdB,cleanmeansE2avCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)}

  # plot an arrow for the final binocular target contrast level that is above 0.8
  if (cond==2){
  arrows(contrastsdB[1:4],abs(cleanmeansE2av[cond,1:4]),contrastsdB[1:4],cleanmeansE2avCI[2,cond,1:4],lwd=3,col=colpal[cond],length=0.05,angle=90)   
  arrows(contrastsdB[5],abs(cleanmeansE2av[cond,5]),contrastsdB[5],0.8,lwd=3,col=colpal[cond],length=0.05,angle=45)    
  }

  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
legend(22,0.8,c('Monocular','Binocular','Dichoptic'),pch=21:23,pt.bg=colpal[1:3],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(f)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF2e2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude (µV) at 4Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

condlist <- c(1,5,6)
pointlist <- c(21,24,25)
for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansE2av[condlist[cond],]), col=colpal[condlist[cond]], lwd=3, cex=0.5)  
  arrows(contrastsdB,abs(cleanmeansE2av[condlist[cond],]),contrastsdB,cleanmeansE2avCI[1,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmeansE2av[condlist[cond],]),contrastsdB,cleanmeansE2avCI[2,condlist[cond],],lwd=3,col=colpal[condlist[cond]],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmeansE2av[condlist[cond],]), pch = pointlist[cond], col='black', bg=colpal[condlist[cond]], cex=1.6, lwd=3)   
}  
legend(22,0.8,c('Monocular','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[condlist],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(g)',adj=0.5,cex=2.5)

dev.off()

postscript("CRF3e2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude (µV) at 3.2Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 4:6){
  lines(contrastsdB,abs(cleanmasksE2av[cond,]), col=colpal[cond], lwd=3, cex=0.5)   
  arrows(contrastsdB,abs(cleanmasksE2av[cond,]),contrastsdB,cleanmasksE2avCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(contrastsdB,abs(cleanmasksE2av[cond,]),contrastsdB,cleanmasksE2avCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  points(contrastsdB,abs(cleanmasksE2av[cond,]), pch = pointlist[cond-3], col='black', bg=colpal[cond], cex=1.6, lwd=3) 
}
legend(22,0.8,c('Monocular 1.6Hz','Binocular cross','Dichoptic cross'),pch=pointlist,pt.bg=colpal[4:6],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(h)',adj=0.5,cex=2.5)

dev.off()

postscript("SpecE.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 8)

frequencies <- (0:299)/10

plotlims <- c(1,6,0,1) 
ticklocsx <- seq(1,6,1)    # locations of tick marks on x axis
ticklocsy <- seq(0,1,0.2)    # locations of tick marks on y axis
ticklabelsx <-ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

polygon(frequencies[c(11:61,61:11)],c(meanspectraECI[1,11:61],meanspectraECI[2,61:11]),col=colpal[2],border=NA)

lines(frequencies[11:61],abs(meanspectraE[11:61]), col=colpal[2], lwd=3, cex=0.5)

text(1.1,0.95,'(b)',adj=0.5,cex=2.5)

dev.off()

postscript("timecourseE.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 8)

times <- seq((1/1000)-1,13,1/1000)
sinewave <- sin(2*times * 2*pi)
sinewave[1:1000] <- 0
sinewave[13001:14000] <- 0

plotlims <- c(-1,13,-10,20) 
ticklocsx <- seq(-1,13,1)    # locations of tick marks on x axis
ticklocsy <- seq(-10,20,10)    # locations of tick marks on y axis
ticklabelsx <- ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

filt <- butter(3,0.083*0.12)
filtwaveE <- signal::filter(filt, meanwavesE)
filtU <- signal::filter(filt,meanwavesECI[1,])
filtL <- signal::filter(filt,meanwavesECI[2,])
polygon(times[c(1:length(times),length(times):1)],c(filtL,filtU[length(filtU):1]),col=colpal[2],border=NA)

lines(times, filtwaveE, col=colpal[2], lwd=1.5)

lines(times, (sinewave*2.5)-7.5, col='black', lwd=3)

text(-0.7,18,'(a)',adj=0.5,cex=2.5)

dev.off()

xpos <- 1:64
ypos <- 1:64
montageE <- toupper(as.character(hdata$Electrode))
for (ch in 1:64){
  i <- match(toupper(electrodes[ch+2]),montageE)
  xpos[ch] <- hdata$X_position[i]
  ypos[ch] <- hdata$Y_position[i]
}

ramp2 <- colorRamp(c("white",colpal[2]))  # create a ramp from one colour to another
colmatrix2 <- rgb(ramp2(seq(0, 1, length = 101)), max = 255)

## Create data frame to be used for interpolation - the function needs columns labelled x, y, and z
rmax <- 0.55   #specify a maximum boundary for the grid
gridRes <- 100 #specify the interpolation grid resolution

toplot <- abs(cleanmeansE[,2,5])
toplot[which(is.na(toplot))] <- 0
testDat<- data.frame(x = xpos, y = -ypos, z = toplot)

#Create the interpolation grid
xo <- seq(min(-rmax, testDat$x), max(rmax, testDat$x), length = gridRes)
yo <- seq(max(rmax, testDat$y), min(-rmax, testDat$y), length = gridRes)

interpV4 <- v4Interp(testDat, xo, yo, rmax, gridRes)

zo2 <- as.matrix(interpV4[,2:ncol(interpV4)])

xo2 <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
yo2 <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
outsidecircle <- sqrt(xo2^2 + yo2^2) > 0.51
zo2[outsidecircle] <- 0

tiff("head1.tiff", height = 600, width = 600, units="px", bg="white")

plotlims <- c(-rmax,rmax,-rmax,rmax) 
par(pty="s")  # make axis square
plot(x=NULL,y=NULL,ann=FALSE, axes=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
image(xo,xo,zo2,col=colmatrix2,zlim=c(0,1),add=TRUE,useRaster=TRUE)
maskx <- c(hdata$OutlineX[1:51]*2.2,hdata$OutlineX[51:1])
masky <- c(hdata$OutlineY[1:51]*2.2,hdata$OutlineY[51:1])
polygon(maskx,masky,border=NA,col="white")
maskx <- c(hdata$OutlineX[51:101]*2.2,hdata$OutlineX[101:51])
masky <- c(hdata$OutlineY[51:101]*2.2,hdata$OutlineY[101:51])
polygon(maskx,masky,border=NA,col="white")

blackelectrodes <- match(toupper(targetelectrodes),toupper(as.character(electrodes[3:66])))
points(xpos[blackelectrodes],ypos[blackelectrodes],pch=16,col="black",cex=2)

lines(hdata$OutlineX,hdata$OutlineY,col="black",lwd=2)
lines(hdata$NoseX,hdata$NoseY,col="black",lwd=2)
lines(hdata$LearX,hdata$LearY,col="black",lwd=2)
lines(hdata$RearX,hdata$RearY,col="black",lwd=2)

dev.off()

toplot <- abs(cleanmeansE2[,2,5])
toplot[which(is.na(toplot))] <- 0
testDat<- data.frame(x = xpos, y = -ypos, z = toplot)

#Create the interpolation grid
xo <- seq(min(-rmax, testDat$x), max(rmax, testDat$x), length = gridRes)
yo <- seq(max(rmax, testDat$y), min(-rmax, testDat$y), length = gridRes)

interpV4 <- v4Interp(testDat, xo, yo, rmax, gridRes)

zo2 <- as.matrix(interpV4[,2:ncol(interpV4)])

xo2 <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
yo2 <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
outsidecircle <- sqrt(xo2^2 + yo2^2) > 0.51
zo2[outsidecircle] <- 0
zo2[which(zo2>1)] <- 1

tiff("head2.tiff", height = 600, width = 600, units="px", bg="white")

plotlims <- c(-rmax,rmax,-rmax,rmax) 
par(pty="s")  # make axis square
plot(x=NULL,y=NULL,ann=FALSE, axes=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
image(xo,xo,zo2,col=colmatrix2,zlim=c(0,1),add=TRUE,useRaster=TRUE)
maskx <- c(hdata$OutlineX[1:51]*2.2,hdata$OutlineX[51:1])
masky <- c(hdata$OutlineY[1:51]*2.2,hdata$OutlineY[51:1])
polygon(maskx,masky,border=NA,col="white")
maskx <- c(hdata$OutlineX[51:101]*2.2,hdata$OutlineX[101:51])
masky <- c(hdata$OutlineY[51:101]*2.2,hdata$OutlineY[101:51])
polygon(maskx,masky,border=NA,col="white")

blackelectrodes <- match(toupper(targetelectrodes),toupper(as.character(electrodes[3:66])))
points(xpos[blackelectrodes],ypos[blackelectrodes],pch=16,col="black",cex=2)

lines(hdata$OutlineX,hdata$OutlineY,col="black",lwd=2)
lines(hdata$NoseX,hdata$NoseY,col="black",lwd=2)
lines(hdata$LearX,hdata$LearY,col="black",lwd=2)
lines(hdata$RearX,hdata$RearY,col="black",lwd=2)

dev.off()

  PostScriptTrace('CRF1e.ps')
  p1 <- readPicture('CRF1e.ps.xml')
  PostScriptTrace('CRF2e.ps')
  p2 <- readPicture('CRF2e.ps.xml')
  PostScriptTrace('CRF3e.ps')
  p3 <- readPicture('CRF3e.ps.xml')
  PostScriptTrace('SpecE.ps')
  p4 <- readPicture('SpecE.ps.xml')
  PostScriptTrace('timecourseE.ps')
  p5 <- readPicture('timecourseE.ps.xml')
  PostScriptTrace('CRF1e2.ps')
  p6 <- readPicture('CRF1e2.ps.xml')
  PostScriptTrace('CRF2e2.ps')
  p7 <- readPicture('CRF2e2.ps.xml')
  PostScriptTrace('CRF3e2.ps')
  p8 <- readPicture('CRF3e2.ps.xml')

    for (n in 1:length(p4@paths)){
    temp <- class(p4@paths[n]$path)[1]
    if (pmatch(temp,"PictureFill",nomatch=0)){
      if (sum(col2rgb(p4@paths[n]$path@rgb))<765){p4@paths[n]$path@rgb <- addalpha(p4@paths[n]$path@rgb,alpha=0.2)}}}

      for (n in 1:length(p5@paths)){
    temp <- class(p5@paths[n]$path)[1]
    if (pmatch(temp,"PictureFill",nomatch=0)){
      if (sum(col2rgb(p5@paths[n]$path@rgb))<765){p5@paths[n]$path@rgb <- addalpha(p5@paths[n]$path@rgb,alpha=0.2)}}}

  h1 <- readTIFF('head1.tiff')
  h2 <- readTIFF('head2.tiff')

  pdf(paste0(figdir,"EEGdata.pdf"), bg="transparent", height = 15, width = 15)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  # insert the head plots first so the white border doesn't overlap the other graphs
    aspratio <- 1  # this is the aspect ratio of the output pdf
  imwidth <- 0.16
  xstart <- 0.65
  ystart <- 0.82
  rasterImage(h1,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth) 
  xstart <- 0.82
  ystart <- 0.82
  rasterImage(h2,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)

  grid.picture(p5,x=0.25,y=0.8,width=0.5,height=1)
  grid.picture(p4,x=0.75,y=0.8,width=0.5,height=1)
  grid.picture(p1,x=0.17,y=0.5,width=0.32,height=1)
  grid.picture(p2,x=0.5,y=0.5,width=0.32,height=1)
  grid.picture(p3,x=0.83,y=0.5,width=0.32,height=1)
  grid.picture(p6,x=0.17,y=0.175,width=0.32,height=1)
  grid.picture(p7,x=0.5,y=0.175,width=0.32,height=1)
  grid.picture(p8,x=0.83,y=0.175,width=0.32,height=1)

    text(0.5,1,'Electroencephalography',cex=4)

  dev.off()

  file.remove(c('CRF1e.ps','CRF2e.ps','CRF3e.ps','CRF1e2.ps','CRF2e2.ps','CRF3e2.ps','SpecE.ps','timecourseE.ps','head1.tiff','head2.tiff'))
  file.remove(c('CRF1e.ps.xml','CRF2e.ps.xml','CRF3e.ps.xml','CRF1e2.ps.xml','CRF2e2.ps.xml','CRF3e2.ps.xml','SpecE.ps.xml','timecourseE.ps.xml'))

}

```

```{r include=FALSE, results='hide'}

# do statistics on the EEG data

alldata <- NULL
groupID <- NULL
allP <- NULL
counter <- 0
for (cont in 1:dim(allmeansP)[3]){
  for (p in 1:dim(allmeansP)[1]){
    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,3,cont])
    groupID[counter] <- cont
    allP[counter] <- p
  }
}
dichE2c <- anovacirc.test(alldata,group=groupID,participant=allP)
# groupID <- as.factor(groupID)
# participantID <- as.factor(participantID)
# adata <- data.frame(dichdata,groupID,participantID)
# dichE2c <- ezANOVA(data = adata, dv = dichdata, within=groupID, wid=participantID)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    for (cond in 1:2){
    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,cond,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- cond
    allP[counter] <- p
    }
  }
}
monbinEstats <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# monbinEstats <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,1,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- 1
    allP[counter] <- p

    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,3,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- 2
    allP[counter] <- p
  }
}
mondichEstats <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# mondichEstats <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,1,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- 1
    allP[counter] <- p

    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,5,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- 2
    allP[counter] <- p
  }
}

monbinEstatsX <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# monbinEstatsX <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

alldata <- NULL
allgroupA <- NULL
allgroupB <- NULL
allP <- NULL
counter <- 0
for (p in 1:30){
  for (lev in 1:5){
    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,1,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- 1
    allP[counter] <- p

    counter <- counter + 1
    alldata[counter] <- mean(allmeansE[p,electrodeindices,6,lev])
    allgroupA[counter] <- lev
    allgroupB[counter] <- 2
    allP[counter] <- p
  }
}

mondichEstatsX <- anovacirc2way(alldata,allgroupA,allgroupB,allP)
# allgroupA <- as.factor(allgroupA)
# allgroupB <- as.factor(allgroupB)
# allP <- as.factor(allP)
# alldata <- abs(alldata)
# adata <- data.frame(alldata,allgroupA,allgroupB,allP)
# mondichEstatsX <- ezANOVA(data = adata, dv = alldata, within=c(allgroupA,allgroupB), wid=allP)

```

```{r EEGdata, fig.cap="Summary of EEG results for N=30 participants. Panel (a) shows a group average waveform for binocular presentation (low pass filtered at 5Hz), with the driving signal plotted at the foot. Panel (b) shows the average Fourier spectrum, and inset scalp distributions. Black dots on the scalp plots indicate electrodes Oz, POz, O1 and O2. Panels (c,d) show contrast response functions at 2Hz for different conditions. Panel (e) shows contrast response functions at 1.6Hz for three conditions. Panels (f-h) are in the same format but for the second harmonic responses. Shaded regions and error bars indicate bootstrapped standard errors.", fig.align="center", echo=FALSE}

knitr::include_graphics(paste0(figdir,'EEGdata.pdf'))

```

When stimuli in both eyes flicker at 2Hz, the binocular responses at the first (Fig. \@ref(fig:EEGdata)c) and second (Fig. \@ref(fig:EEGdata)f) harmonics are substantially greater than the monocular responses, particularly at high contrasts. Analysis of variance on the complex values ($ANOVA^2_{circ}$) revealed a main effect of contrast (F(`r monbinEstats$dfA`,`r monbinEstats$dfR`) = `r round(monbinEstats$FratioA,digits=2)`, $p$ < 0.001) and an interaction effect (F(`r monbinEstats$dfAB`,`r monbinEstats$dfR`) = `r round(monbinEstats$FratioAB,digits=2)`, $p$ < 0.001), but no effect of condition ($p$ = `r round(monbinEstats$pvalB,digits=2)`) at the first harmonic, with a similar pattern of results obtained at the second harmonic (note that the absence of a main effect of condition is expected, given that there should be negligible differences between conditions at low contrasts, and any differences are therefore captured by the interaction term). For the cross-frequency conditions (Fig. \@ref(fig:EEGdata)d,g), there was no appreciable effect of adding a 1.6Hz component on the response at 2Hz or 4Hz (no effect of condition, and no interaction). Similarly, there were no clear interocular interactions between frequencies in the responses at 1.6Hz (Fig. \@ref(fig:EEGdata)e) and 3.2Hz (Fig. \@ref(fig:EEGdata)h). This pattern of results suggests that processing of temporal luminance modulations happens in a more linear way in visual cortex (indexed by EEG), compared with subcortical pathways (indexed by pupillometry), and shows no evidence of interocular suppression.

```{r include=FALSE, results='hide'}

# generate plot showing the bin:mon ratio for three data sets

         if (!file.exists(paste0(datadir,'Averagedata.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='Averagedata.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }

load(file=paste0(datadir,'Averagedata.RData'))

contrastsdB <- 20*log10(c(6,12,24,48,96))
electrodeindices <- match(targetelectrodes,electrodes)-2

allPratios <- 20*log10(abs(allmeansP[,2,])/abs(allmeansP[,1,]))
allEratios <- 20*log10(abs(apply(allmeansE[,electrodeindices,2,],c(1,3),mean))/abs(apply(allmeansE[,electrodeindices,1,],c(1,3),mean)))
allE2ratios <- 20*log10(abs(apply(allmeansE2[,electrodeindices,2,],c(1,3),mean))/abs(apply(allmeansE2[,electrodeindices,1,],c(1,3),mean)))

allratios <- NULL
allconds <- NULL
allconts <- NULL
allparticipants <- NULL
counter <- 0
for (cont in 1:length(contrastsdB)){
  for (s in 1:nrow(allPratios)){
    counter <- counter + 1
    allratios[counter] <- allPratios[s,cont]
    allconds[counter] <- 1
    allparticipants[counter] <- s
    allconts[counter] <- cont

    counter <- counter + 1
    allratios[counter] <- allEratios[s,cont]
    allconds[counter] <- 2
    allparticipants[counter] <- s
    allconts[counter] <- cont

    counter <- counter + 1
    allratios[counter] <- allE2ratios[s,cont]
    allconds[counter] <- 3
    allparticipants[counter] <- s
    allconts[counter] <- cont

  }
}
allparticipants <- as.factor(allparticipants)
allconds <- as.factor(allconds)
allconts <- as.factor(allconts)
aovdata <- data.frame(allparticipants,allconds,allconts,allratios)
ratioaov <- ezANOVA(aovdata,dv=allratios,within=c(allconds,allconts),wid=allparticipants)
gg <- ratioaov$`Sphericity Corrections`$GGe

if (processdata > 0){

PratiosCI <- matrix(0,nrow=2,ncol=5)
EratiosCI <- matrix(0,nrow=2,ncol=5)
E2ratiosCI <- matrix(0,nrow=2,ncol=5)
for (level in 1:5){
    bspop <- NULL
    for (n in 1:nbootstraps){
      bspop[n] <- mean(sample(allPratios[,level],nrow(allPratios),replace=TRUE))
    }
    PratiosCI[,level] <- quantile(bspop,c((1-0.68)/2,1-((1-0.68)/2)))
     bspop <- NULL
    for (n in 1:nbootstraps){
      bspop[n] <- mean(sample(allEratios[,level],nrow(allEratios),replace=TRUE))
    }
    EratiosCI[,level] <- quantile(bspop,c((1-0.68)/2,1-((1-0.68)/2)))
    bspop <- NULL
    for (n in 1:nbootstraps){
      bspop[n] <- mean(sample(allE2ratios[,level],nrow(allE2ratios),replace=TRUE))
    }
    E2ratiosCI[,level] <- quantile(bspop,c((1-0.68)/2,1-((1-0.68)/2)))
}

plotlims <- c(15,40,-3,12)  
ticklocsx <- contrastsdB
ticklabelsx <-c(6,12,24,48,96) 
ticklocsy <- seq(-3,12,3)    # locations of tick marks on y axis
ticklabelsy <- c(expression(1/sqrt(2)),'1',expression(sqrt(2)),'2',expression(2*sqrt(2)),'4')

pdf("Figures/BSratios.pdf", onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Bin:Mon ratio", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

arrows(contrastsdB,colMeans(allPratios),contrastsdB,PratiosCI[1,],lwd=3,col=colpal[3],length=0.05,angle=90)
arrows(contrastsdB,colMeans(allPratios),contrastsdB,PratiosCI[2,],lwd=3,col=colpal[3],length=0.05,angle=90)
arrows(contrastsdB,colMeans(allEratios),contrastsdB,EratiosCI[1,],lwd=3,col=colpal[4],length=0.05,angle=90)
arrows(contrastsdB,colMeans(allEratios),contrastsdB,EratiosCI[2,],lwd=3,col=colpal[4],length=0.05,angle=90)
arrows(contrastsdB,colMeans(allE2ratios),contrastsdB,E2ratiosCI[1,],lwd=3,col=colpal[5],length=0.05,angle=90)
arrows(contrastsdB,colMeans(allE2ratios),contrastsdB,E2ratiosCI[2,],lwd=3,col=colpal[5],length=0.05,angle=90)

lines(contrastsdB,colMeans(allPratios), col=colpal[3], lwd=3, cex=0.5)
lines(contrastsdB,colMeans(allEratios), col=colpal[4], lwd=3, cex=0.5)
lines(contrastsdB,colMeans(allE2ratios), col=colpal[5], lwd=3, cex=0.5)

points(contrastsdB,colMeans(allPratios), pch = 21, col='black', bg=colpal[3], cex=1.6, lwd=3) 
points(contrastsdB,colMeans(allEratios), pch = 22, col='black', bg=colpal[4], cex=1.6, lwd=3)
points(contrastsdB,colMeans(allE2ratios), pch = 23, col='black', bg=colpal[5], cex=1.6, lwd=3)

legend(15,12,c('Pupillometry','EEG 1F', 'EEG 2F'), pch=21:23, pt.bg=colpal[3:5],box.lwd=2,pt.cex=1.6,pt.lwd=3)

dev.off()

}

```

Finally, we calculated the ratio of binocular to monocular responses across the three data types from Experiment 1. Fig. \@ref(fig:BSratios) shows that these ratios are approximately $\sqrt2$ across the low-to-intermediate contrast range for all three data types. At higher contrasts, we see ratios of 2 or higher for the EEG data, but much weaker ratios near 1 for the pupillometry data. Note that the ratios here are calculated on a per-participant basis and then averaged, rather than being the ratios of the average values shown in Fig. \@ref(fig:pupildata) and \@ref(fig:EEGdata). A $3 \times 5$ repeated measures ANOVA on the logarithmic (dB) ratios found a main effect of contrast (F(`r round(ratioaov$ANOVA$DFn[2]*gg[2],digits=2)`,`r round(ratioaov$ANOVA$DFd[2]*gg[2],digits=2)`) = `r round(ratioaov$ANOVA$F[2],digits=2)`, \emph{p} < `r round(ratioaov$ANOVA$p[2],digits=3)`), no effect of data modality (F(`r ratioaov$ANOVA$DFn[1]`,`r ratioaov$ANOVA$DFd[1]`) = `r round(ratioaov$ANOVA$F[1],digits=2)`, \emph{p} = `r round(ratioaov$ANOVA$p[1],digits=2)`), but a highly significant interaction (F(`r round(ratioaov$ANOVA$DFn[3]*gg[3],digits=2)`,`r round(ratioaov$ANOVA$DFd[3]*gg[3],digits=2)`) = `r round(ratioaov$ANOVA$F[3],digits=2)`, \emph{p} < `r round(ratioaov$ANOVA$p[3],digits=4)`).

```{r BSratios, fig.cap="Ratio of binocular to monocular response for three data types. Each ratio is the average of ratios for N=30 participants, and error bars indicate bootstrapped standard errors.", fig.align="center", out.width = '50%', echo=FALSE}

knitr::include_graphics(paste0(figdir,'BSratios.pdf'))

```

## Experiment 2

The strong binocular facilitation and weak interocular suppression in the EEG data from Experiment 1 was very different from previous findings on binocular combination using steady-state EEG with grating stimuli [@Baker2017]. One possible explanation is that the lower temporal frequency used here (2Hz, vs 5 or 7Hz in previous work) might be responsible for this difference. We therefore ran a second experiment to compare monocular and binocular responses at a range of temporal frequencies. Only EEG data were collected for this experiment, as the pupil response is negligible above around 2Hz [@Spitschan2014]; note that we originally chose 2Hz because it produces measurable signals for both EEG and pupillometry, yet is unfortunately optimal for neither.

```{r include=FALSE, results='hide'}

# (download and) analyse the raw data from the temporal frequency experiment

if (processdata > 2){

  legaltriggers <- 1

  TFs <- c(2,4,8,16,30)
  EEGtimes <- seq(1/1000,10,1/1000)
  targetindices <- (TFs*10)+1

  for (participant in 1:12){
    print(participant)

# download and process data if this has not been done already for this participant
    if (!file.exists(paste0(datadir,'P',150+participant,'summary.RData'))){

      # if there is no directory for this participant's data
      if (!file.exists(paste0(rawdir,'P',150+participant))){
        dir.create(paste0(rawdir,'P',150+participant))}

      d <- dir(paste0(rawdir,'P',150+participant))

      if (length(d)==0){

        # if the tar file of this participant's data doesn't exist
        if (!file.exists(paste0(rawdir,'P',150+participant,'.tar'))){
          # download it from OSF
          if (!exists('tffiles')){
            osfnode <- 'mvpyz'
            osfproject <- osf_retrieve_node(osfnode)
            tffiles <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(tffiles$name==paste0('P',150+participant,'.tar'))
          osf_download(tffiles[fid,],rawdir,progress=TRUE)
        }

        # then unzip the tar file
        untar(paste0(rawdir,'P',150+participant,'.tar'),exdir=paste0(rawdir,'P',150+participant))

        # and delete the tar file to save storage space
        file.remove(paste0(rawdir,'P',150+participant,'.tar'))
      }

      d <- dir(paste0(rawdir,'P',150+participant),full.names=TRUE)

      EEGtargets <- array(0,dim=c(5,64,20,5))
      EEGspectra <- array(0,dim=c(5,64,20,500))

      psychopyfiles <- dir(path=paste0(rawdir,'P',150+participant), pattern=paste0('P',150+participant,'_CRFstudy'), full.names = TRUE)

      EEGfiles <- dir(path=paste0(rawdir,'P',150+participant), pattern='*_EEG.gz', full.names = TRUE)

      for (block in 1:length(EEGfiles)){

        psychoutput <- read.csv(psychopyfiles[block])

        condorder <- psychoutput$condition

        EEGdata <- read.csv(EEGfiles[block])
        electrodes <- colnames(EEGdata)

        triggertimes <- NULL
        counter <- 0
        lasttrigger <- -10000
        for (n in 1:nrow(EEGdata)){
          if(EEGdata$Trigger[n] %in% legaltriggers){
            if(n>(lasttrigger+10000)){
              counter <- counter + 1
              triggertimes[counter] <- n
              lasttrigger <- n
            }}
        }

        for (cond in 1:counter){
          for (ch in 1:64){
            trial <- EEGdata[(triggertimes[cond]+(2001:12000)),ch+2]
            fspec <- (fft(trial)/length(trial))
            EEGtargets[block,ch,condorder[cond],] <- fspec[targetindices]
            EEGspectra[block,ch,condorder[cond],1:500] <- fspec[1:500]
          }}
      }

      electrodeindices <- match(targetelectrodes,electrodes)-2

      cleanmeansE <- array(0,dim=c(2,5))
      meanspec <- array(0,dim=c(2,5,500))

      for (eye in 1:2){
        for (cond in 1:5){
          includedconds <- c(cond,cond+5) + (10*(eye-1))
          temp <- as.vector(EEGtargets[,electrodeindices,includedconds,cond])
          temp <- temp[which(!is.na(temp))]
          temp <- temp[which(abs(temp)>0)]  
          cartdata <- data.frame(Re(temp),Im(temp))
          if (sum(cartdata)!=0){
            D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
            i <- which(D<SDthresh) 
            cleanmeansE[eye,cond] <- mean(temp[i])}

          for (f in 2:500){
            temp <- as.vector(apply(EEGspectra[,electrodeindices,includedconds,f],c(1,3),mean))
            temp <- temp[which(!is.na(temp))]
            temp <- temp[which(abs(temp)>0)]  
            cartdata <- data.frame(Re(temp),Im(temp))
            if (sum(cartdata)!=0){
              D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
              i <- which(D<SDthresh) 
              meanspec[eye,cond,f] <- mean(temp[i])}   
          }
        }
      }

    }

    save(file=paste0(datadir,'P',150+participant,'summary.RData'),list=c('cleanmeansE','meanspec'))

  }

}

```

```{r include=FALSE, results='hide'}

# load in the processed data from the temporal frequency experiment and average

if (processdata > 1){

  allspec <- array(0,dim=c(12,2,5,500))
  allmeans <- array(0,dim=c(12,2,5))
  allratios <- matrix(0,nrow=12,ncol=5)

      for (participant in 1:12){
         # if the summary file of this participant's data doesn't exist
         if (!file.exists(paste0(datadir,'P',150+participant,'summary.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name==paste0('P',150+participant,'summary.RData'))
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }
      }

  for (participant in 1:12){

    load(paste0(datadir,'P',150+participant,'summary.RData'))

    allmeans[participant,,] <- cleanmeansE
    allspec[participant,,,] <- meanspec
    allratios[participant,] <- 20*log10(abs(allmeans[participant,2,])/abs(allmeans[participant,1,]))
  }

meanratios <- NULL
meanratiosCI <- matrix(0,nrow=2,ncol=5)
cleanmeansTF <- matrix(0,nrow=2,ncol=5)
cleanmeansTFCI <- array(0,dim=c(2,2,5))
meanspectraTF <- array(0,dim=c(2,5,500))
meanspectraTFCI <- array(0,dim=c(2,2,5,500))

for (cond in 1:2){
  for (level in 1:5){

    temp <- allmeans[,cond,level]
    cartdata <- data.frame(Re(temp),Im(temp))
    if (sum(cartdata)!=0){
      D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
      i <- which(D<SDthresh) 
    mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
    cleanmeansTF[cond,level] <- mbs$meanamp
    cleanmeansTFCI[1,cond,level] <- mbs$lowerCI
    cleanmeansTFCI[2,cond,level] <- mbs$upperCI
    }

    meanratios[level] <- mean(allratios[,level])
    bspop <- NULL
    for (n in 1:nbootstraps){
      bspop[n] <- mean(sample(allratios[,level],nrow(allratios),replace=TRUE))
    }
    meanratiosCI[,level] <- quantile(bspop,c((1-0.68)/2,1-((1-0.68)/2)))

    for (f in 2:500){

        temp <- allspec[,cond,level,f]
      cartdata <- data.frame(Re(temp),Im(temp))
      if (sum(cartdata)!=0){
        D <- sqrt(stats::mahalanobis(cartdata, colMeans(cartdata), cov(cartdata)))
        i <- which(D<SDthresh) 
mbs <- amperrors(temp[i],method='boot',quantiles=68,nresamples=nbootstraps)
meanspectraTF[cond,level,f] <- mbs$meanamp
meanspectraTFCI[1,cond,level,f] <- mbs$lowerCI
meanspectraTFCI[2,cond,level,f] <- mbs$upperCI
      }
    }
  }
}

save(file=paste0(datadir,'AveragedataTF.RData'),list=c('cleanmeansTF','cleanmeansTFCI','meanspectraTF','meanspectraTFCI','meanratios','meanratiosCI'))

}

```

```{r include=FALSE, results='hide'}

# generate figure showing the TF data

if (processdata > 0){

         if (!file.exists(paste0(datadir,'AveragedataTF.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='AveragedataTF.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }

  load(paste0(datadir,'AveragedataTF.RData'))

  TFs <- c(2,4,8,16,30)
  TFsdB <- 20*log10(TFs)

  postscript("SpecTF.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 8, width = 5.5)

frequencies <- (0:499)/10

plotlims <- c(0,36,0,5) 
ticklocsx <- seq(0,36,6)    # locations of tick marks on x axis
ticklocsy <- seq(0,5,1)    # locations of tick marks on y axis
ticklabelsx <- 2^(0:6)        # set labels for x ticks
ticklabelsy <- c(0,1,'','','','')    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)

for (cond in 1:5){

lines(c(0,36),c(cond,cond)-1)  
polygon(20*log10(frequencies[c(11:490,490:11)]),(cond-1)+c(meanspectraTFCI[1,2,cond,11:490],meanspectraTFCI[2,2,cond,490:11]),col=colpal[cond],border=NA)

lines(20*log10(frequencies[11:490]),(cond-1)+abs(meanspectraTF[2,cond,11:490]), col=colpal[cond], lwd=2, cex=0.5) 
}

text(33,0.25,'2Hz',adj=0.5,cex=1.5,col=colpal[1])
text(33,1.25,'4Hz',adj=0.5,cex=1.5,col=colpal[2])
text(33,2.25,'8Hz',adj=0.5,cex=1.5,col=colpal[3])
text(33,3.25,'16Hz',adj=0.5,cex=1.5,col=colpal[4])
text(33,4.25,'30Hz',adj=0.5,cex=1.5,col=colpal[5])

text(1.5,4.8,'(a)',adj=0.5,cex=2.5)
text(18,4.8,'Binocular',adj=0.5,cex=2)

dev.off()

plotlims <- c(6,30,0,1)  
ticklocsx <- seq(6,30,6)    # locations of tick marks on x axis
ticklocsy <- seq(0,1,0.2)    # locations of tick marks on y axis
ticklabelsx <- 2^(1:5)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("TF1.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:2){
  lines(TFsdB,abs(cleanmeansTF[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  arrows(TFsdB,abs(cleanmeansTF[cond,]),TFsdB,cleanmeansTFCI[1,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  arrows(TFsdB,abs(cleanmeansTF[cond,]),TFsdB,cleanmeansTFCI[2,cond,],lwd=3,col=colpal[cond],length=0.05,angle=90)
  points(TFsdB,abs(cleanmeansTF[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
legend(20,1,c('Monocular','Binocular'),pch=21:22,pt.bg=colpal[1:2],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(7.5,0.92,'(b)',adj=0.5,cex=2.5)

dev.off()

plotlims <- c(6,30,-3,12)  
ticklocsy <- seq(-3,12,3)    # locations of tick marks on y axis
ticklabelsy <- c(expression(1/sqrt(2)),'1',expression(sqrt(2)),'2',expression(2*sqrt(2)),'4')

postscript("TF2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Frequency (Hz)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Bin:Mon ratio", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

  lines(TFsdB,meanratios, col=colpal[3], lwd=3, cex=0.5)
  arrows(TFsdB,meanratios,TFsdB,meanratiosCI[1,],lwd=3,col=colpal[3],length=0.05,angle=90)
  arrows(TFsdB,meanratios,TFsdB,meanratiosCI[2,],lwd=3,col=colpal[3],length=0.05,angle=90)
  points(TFsdB,meanratios, pch = 20+3, col='black', bg=colpal[3], cex=1.6, lwd=3)   
 text(7.5,11,'(c)',adj=0.5,cex=2.5)

dev.off()

  PostScriptTrace('SpecTF.ps')
  p1 <- readPicture('SpecTF.ps.xml')
  PostScriptTrace('TF1.ps')
  p2 <- readPicture('TF1.ps.xml')
  PostScriptTrace('TF2.ps')
  p3 <- readPicture('TF2.ps.xml')

    for (n in 1:length(p1@paths)){
    temp <- class(p1@paths[n]$path)[1]
    if (pmatch(temp,"PictureFill",nomatch=0)){
      if (sum(col2rgb(p1@paths[n]$path@rgb))<765){p1@paths[n]$path@rgb <- addalpha(p1@paths[n]$path@rgb,alpha=0.2)}}}

  pdf(paste0(figdir,"TFdata.pdf"), bg="transparent", height = 12, width = 15)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  grid.picture(p1,x=0.3,y=0.5,width=0.6,height=1)
  grid.picture(p2,x=0.78,y=0.74,width=0.4,height=1)
  grid.picture(p3,x=0.78,y=0.26,width=0.4,height=1)

  dev.off()

  file.remove(c('TF1.ps','TF2.ps','SpecTF.ps','TF1.ps.xml','TF2.ps.xml','SpecTF.ps.xml'))

}

```

Results from the temporal frequency experiment are shown in Fig. \@ref(fig:TFdata). Fig. \@ref(fig:TFdata)a shows the Fourier spectra for responses to binocular flicker at 5 different frequencies (2, 4, 8, 16, and 30 Hz). From 2 to 16 Hz, clear signals are observed at each fundamental frequency, and typically also their higher harmonics (integer multiples of the fundamental). However, at 30 Hz (upper row), the responses recorded were not demonstrably above the noise baseline. Fig. \@ref(fig:TFdata)b compares the monocular and binocular responses at each stimulation frequency. Here we replicate the substantial summation effect across frequencies up to and including 16Hz (Fig. \@ref(fig:TFdata)c), demonstrating that strong binocular facilitation in the EEG data of Experiment 1 cannot be attributed to our use of 2Hz flicker.

```{r TFdata, fig.cap="Binocular facilitation at different temporal frequencies. Panel (a) shows Fourier spectra for responses to binocular flicker at 5 different frequencies (offset vertically for clarity). Panel (b) shows the response at each stimulation frequency for monocular (red circles) and binocular (blue squares) presentation. Panel (c) shows the ratio of binocular to monocular responses. Error bars and shaded regions indicate bootstrapped standard errors across N=12 participants.", fig.align="center", echo=FALSE}

knitr::include_graphics(paste0(figdir,'TFdata.pdf'))

```

## Experiment 3

In Experiment 1 we found evidence of stronger binocular facilitation for cortical responses to luminance flicker (measured using EEG), compared with subcortical responses (measured using pupillometry). Since perception is dependent on cortical responses, these results provide a clear prediction for perceived contrast judgements indexed by psychophysical contrast matching paradigms [@Anstis1998; @Levelt1965; @Quaia2018]. We therefore conducted such an experiment, in which participants judged which of two stimuli had the greater perceived amplitude of flicker. One stimulus was a matching stimulus, that had a fixed binocular flicker amplitude of either 24\% or 48\% (temporal) contrast. The other stimulus was a target stimulus, the contrast of which was controlled by a staircase algorithm. We tested 9 ratios of contrast between the left and right eyes.

The results from the matching experiment are shown in Fig. \@ref(fig:matchingdata). Each data point indicates the contrast levels required in each eye that were perceptually equivalent to the binocular 24\% (red circles) and 48\% (blue circles) matching contrasts. At both matching contrasts, we see a very substantial increase in the physical contrast required for a monocular target (data points along the x- and y-axes), compared to a binocular target (points along the diagonal of x=y). For example with a 48\% match, the monocular targets required contrasts close to 100\%, whereas binocular targets required a contrast of around 50\%. The data points between these extremes also fall close to the predictions of a linear summation model (diagonal dotted lines), and are inconsistent with a winner-takes-all (or MAX) model (dashed lines). Overall, these matching results are consistent with the approximately linear summation effects observed in the EEG data of Experiment 1 (Fig. \@ref(fig:EEGdata)c,f).

```{r include=FALSE, results='hide'}

# (download and) analyse individual participant data for the matching experiment

ratiosL <- c(1, 1, 1, 1, 1, 0.75, 0.5, 0.25, 0)
ratiosR <- c(0, 0.25, 0.5, 0.75, 1, 1, 1, 1, 1)

if (processdata > 2){

if (!file.exists(paste0(datadir,'matchingdata.csv'))){
osfnode <- 'ey7rg'  
osfproject <- osf_retrieve_node(osfnode)
matchingfiles <- osf_ls_files(osfproject,n_max=300)
fid <- which(matchingfiles$name=='matchingdata.csv')
osf_download(matchingfiles[fid,],datadir,progress=TRUE)
}

matchingdata <- read.csv(paste0(datadir,'matchingdata.csv'))

levels <- unique(matchingdata$allref)
subjects <- unique(matchingdata$Participant)

allthresh <- array(0,dim=c(length(subjects),length(levels),length(ratiosL)))
for (subj in 1:length(subjects)){
  thissubj <- subset(matchingdata,Participant==subj)  
  levels <- unique(thissubj$allref)
for (cond in 1:length(levels)){
  thiscond <- subset(thissubj,allref==levels[cond])
  contrast <- levels[cond]
  for (ll in 1:length(ratiosL)){
    Lratio <- ratiosL[ll]
    Rratio <- ratiosR[ll]
    thisL <- subset(thiscond, allLratio==Lratio)
    thisdata <- subset(thisL, allRratio==Rratio)

Ncorrect <- NULL
Ntrials <- NULL
Clevs <- sort(unique(thisdata$allcontrast))
for (lev in 1:length(Clevs)){
  temp <- subset(thisdata,allcontrast==Clevs[lev])
  Ncorrect[lev] <- sum(temp$correct)
  Ntrials[lev] <- nrow(temp)
}
CdB <- 20*log10(Clevs)
datatofit <- data.frame(CdB,Ncorrect,Ntrials)
fit <- quickpsy::quickpsy(datatofit, x=CdB, k=Ncorrect, n=Ntrials, guess=0, xmin=0, xmax=40, B = 0, fun=cum_normal_fun)
allthresh[subj,cond,ll] <- fit$thresholds$thre

# plot(datatofit$CdB,datatofit$Ncorrect/datatofit$Ntrials)
# points(datatofit$CdB,datatofit$Ncorrect/datatofit$Ntrials,pch=16)
# lines(fit$curves$x,fit$curves$y)

  }
}
}

allthresh[1,2,] <- NA   # exclude outliers
allthresh[2,2,2] <- NA  # exclude outliers
meanthresh <- apply(allthresh,2:3,mean,na.rm=TRUE)
SEthresh <- apply(allthresh,2:3,sd,na.rm=TRUE)/sqrt(apply(!is.na(allthresh),2:3,sum))

save(file=paste0(datadir,'meanmatching.RData'),list=c('meanthresh','SEthresh','allthresh'))

}

if (processdata > 0){

           if (!file.exists(paste0(datadir,'meanmatching.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='meanmatching.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
           }

load(paste0(datadir,'meanmatching.RData'))

pdf(paste0(figdir,"matchingdata.pdf"), bg="transparent", height = 5.5, width = 5.5)

plotlims <- c(0,125,0,125)  
ticklocs <- seq(0,125,25)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs)     # add the tick labels
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="Right eye contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Left eye contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

lines(c(0,24),c(24,24),lty=2)
lines(c(0,48),c(48,48),lty=2)
lines(c(24,24),c(0,24),lty=2)
lines(c(48,48),c(0,48),lty=2)
lines(c(0,96),c(96,0),lty=3)
lines(c(0,48),c(48,0),lty=3)

for (matchlev in 1:2){
lines(ratiosR*10^(meanthresh[matchlev,]/20),ratiosL*10^(meanthresh[matchlev,]/20),pch=16,col=colpal[matchlev],lwd=3)}

# # these lines of code plot the individual participants' data for checking
# for (subj in 1:10){
# for (matchlev in 1:2){
#   lines(ratiosL*10^(allthresh[subj,matchlev,]/20),ratiosR*10^(allthresh[subj,matchlev,]/20),pch=16,col=colpal[matchlev],lwd=1)
# for (cond in 1:9){
#   points(ratiosL[cond]*10^(allthresh[subj,matchlev,cond]/20),ratiosR[cond]*10^(allthresh[subj,matchlev,cond]/20),pch=16,cex=1,col=colpal[matchlev])}}}

for (matchlev in 1:2){
for (cond in 1:9){
  points(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),pch=16,cex=2,col=colpal[matchlev])

  if ((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])<42){
  arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),ratiosR[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),ratiosL[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),length=0.015, angle=90, lwd=2)
  }

  # plot as an arrow if exceeding the axis limits
  if ((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])>42){
  arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),125,ratiosL[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),length=0.035, angle=45, lwd=2)
  }

  arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),ratiosR[cond]*10^((meanthresh[matchlev,cond]-SEthresh[matchlev,cond])/20),ratiosL[cond]*10^((meanthresh[matchlev,cond]-SEthresh[matchlev,cond])/20),length=0.015, angle=90, lwd=2)
}
}

legend(75,95,c('48% match','24% match'),pch=16,col=colpal[2:1],pt.cex=2,box.lwd=2)

legend(45.75,125,c('Winner-take-all','Linear summation'),lty=2:3,box.lwd=2)

dev.off()

}

```

```{r matchingdata, fig.cap="Contrast matching functions. Dotted and dashed lines are predictions of canonical summation models with a linear exponent (dotted) or an infinite exponent (dashed). Error bars indicate the standard error across participants (N=10), and are constrained along radial lines converging at the origin. Note that, for the {48\\%} match, the data point on the x axis falls higher than {100\\%} contrast. This is because the psychometric function fits for some individuals were interpolated such that the PSE fell above {100\\%}, shifting the mean slightly above that value.", out.width = '50%', fig.align="center", echo=FALSE}

knitr::include_graphics(paste0(figdir,'matchingdata.pdf'))

```

## Computational modelling

We fitted a computational model to the data from Experiments 1 & 3 using a hierarchical Bayesian approach. The model behaviour is displayed in Fig. \@ref(fig:modelfigure)e-h, with empirical data replotted in Fig. \@ref(fig:modelfigure)a-d for comparison. In general, the model captures the key characteristics of the empirical data. Note that there are some minor discrepancies, which are a consequence of the hierarchical nature of the modelling. In brief, the model is fitted to the amplitudes for each participant, and group-level parameter estimates are derived based on these fits (see Table \@ref(tab:paramtable)). This procedure discards the phase information, whereas the empirical averages are coherently averaged across participants (retaining phase information). This explains the amplitude differences between model and data, particularly at low target contrast levels, but is of little consequence for the pattern of relative responses across conditions, which is our main focus here.

We were particularly interested in comparing the weight of interocular suppression across data sets. We therefore plot the posterior distributions for this parameter for all four data sets (see Fig. \@ref(fig:modelfigure)i). The key finding is that the pupillometry results (green distribution) display a much greater weight of interocular suppression compared with the other data sets (grey, purple and yellow distributions). There is no overlap between the pupillometry distribution and any of the other three. All four distributions are also meaningfully below a weight of 1 -- the value that previous work using grating stimuli would predict [@Meese2006; @Baker2017], and the peak location of our prior distribution (black curve). These results offer an explanation of the empirical data: the strong interocular suppression for the pupillometry data is consistent with the weak binocular facilitation, and significant dichoptic masking observed using that method. The weaker suppression for the other experiments is consistent with the near-linear binocular facilitation effects, and absent dichoptic masking.

```{r include=FALSE, results='hide'}

# create text objects for Stan models

CRFmodelstring = "
  data {
    int<lower=1> Nsubj ;
    int<lower=1> Ntotal ;
    real y[Ntotal] ;
    real x[Ntotal] ;
    real<lower=0> w[Ntotal] ;
    int<lower=1> c[Ntotal] ;
    int<lower=1> s[Ntotal] ;
  }
  parameters {
    real<lower=1> Z[Nsubj] ;
    real<lower=0> Rmax[Nsubj] ;
    real<lower=0> wt[Nsubj] ;
    real<lower=0> k[Nsubj] ;
    real<lower=0> sigma ;
    real<lower=1> Zmu ; 
    real<lower=0> Rmaxmu ; 
    real<lower=0> wtmu ; 
    real<lower=0> kmu ; 
    real<lower=0> Zsigma ;
    real<lower=0> Rmaxsigma ;
    real<lower=0> wtsigma ;
    real<lower=0> ksigma ;
    real<lower=0> nu ;
  }
  model {
    Zmu ~ normal( 5 , 2 ) ;
    Rmaxmu ~ normal( 1 , 0.5 ) ;
    wtmu ~ normal( 1 , 0.5 ) ;
    kmu ~ normal( 5 , 2 ) ;
    sigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    Zsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    Rmaxsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    wtsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    ksigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    nu ~ exponential(1/30.0) ;
    Z ~ normal( Zmu , Zsigma ) ; // vectorized
    Rmax ~ normal( Rmaxmu , Rmaxsigma ) ; // vectorized
    wt ~ normal( wtmu , wtsigma ) ; // vectorized
    k ~ normal( kmu , ksigma ) ; // vectorized
    for ( i in 1:Ntotal ) {

    if (c[i]==1)  // monocular condition
       y[i] ~ student_t( 
                nu ,
                k[s[i]] + (Rmax[s[i]]*(pow(x[i],2))/(Z[s[i]] + x[i])) ,
                w[i]*sigma ) ;

     else if (c[i]==2)  // binocular condition 
       y[i] ~ student_t( 
                nu ,
                k[s[i]] + (Rmax[s[i]]*((pow(x[i],2))/(Z[s[i]] + x[i] + wt[s[i]]*x[i]) + (pow(x[i],2))/(Z[s[i]] + x[i] + wt[s[i]]*x[i]))),
                w[i]*sigma ) ;

     else if (c[i]==3)  // dichoptic condition
       y[i] ~ student_t( 
                nu ,
                k[s[i]] + (Rmax[s[i]]*((pow(x[i],2))/(Z[s[i]] + x[i] + wt[s[i]]*48) + (pow(48,2))/(Z[s[i]] + 48 + wt[s[i]]*x[i]))),
                w[i]*sigma ) ;

     else if (c[i]==4)  // binocular cross condition 
       y[i] ~ student_t( 
                nu ,
                k[s[i]] + (Rmax[s[i]]*(pow(x[i],2))/(Z[s[i]] + x[i] + wt[s[i]]*x[i])) ,
                w[i]*sigma ) ;

     else if (c[i]==5)  // dichoptic cross condition
       y[i] ~ student_t( 
                nu ,
                k[s[i]] + (Rmax[s[i]]*(pow(x[i],2))/(Z[s[i]] + x[i] + wt[s[i]]*48)) ,
                w[i]*sigma ) ;       
    }
  }

  " # close quote for CRFmodelstring

matchingmodelstring = "
  data {
    int<lower=1> Nsubj ;
    int<lower=1> Ntotal ;
    int correct[Ntotal] ;
    real allcontrast[Ntotal] ;
    real allref[Ntotal] ;
    real allLratio[Ntotal] ;
    real allRratio[Ntotal] ;    
    int<lower=1> s[Ntotal] ;
  }
  parameters {
    real<lower=0> Z[Nsubj] ;
    real<lower=0> k[Nsubj] ;
    real<lower=0> w[Nsubj] ;
    real<lower=0> sigma ;
    real<lower=0> Zmu ; 
    real<lower=0> kmu ; 
    real<lower=0> wmu ;     
    real<lower=0> Zsigma ;
    real<lower=0> ksigma ;
    real<lower=0> wsigma ;
    }
  model {
    Zmu ~ normal( 5 , 2 ) ;
    wmu ~ normal( 1 , 0.5 ) ;
    kmu ~ normal( 5 , 2 ) ;
    sigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    Zsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    ksigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    wsigma ~ uniform( 1.0E-3 , 1.0E+3 ) ;
    Z ~ normal( Zmu , Zsigma ) ; // vectorized
    k ~ normal( kmu , ksigma ) ; // vectorized
    w ~ normal( wmu , wsigma ) ; // vectorized

    for ( i in 1:Ntotal ) {

      real standresp ;
      real matchrespL ;
      real matchrespR ;
      real matchresp ;

      standresp = 2 * pow(allref[i],2) / (Z[s[i]] + allref[i] + w[s[i]]*allref[i]);
      matchrespL = pow(allLratio[i]*allcontrast[i],2) / (Z[s[i]] + allLratio[i]*allcontrast[i] + w[s[i]]*allRratio[i]*allcontrast[i]);
      matchrespR = pow(allRratio[i]*allcontrast[i],2) / (Z[s[i]] + allRratio[i]*allcontrast[i] + w[s[i]]*allLratio[i]*allcontrast[i]);
      matchresp = matchrespL + matchrespR;
      correct[i] ~ bernoulli(((1 + erf((matchresp - standresp)/k[s[i]]))/2)*0.998 + 0.001) ;
    }
  }  
  " # close quote for matchingmodelstring

```

```{r include=FALSE, results='hide'}

# do computational modelling

if (processdata > 1){

# settings for sampler 
numSavedSteps <- 100000    # total number of MCMC steps
adaptSteps <- 1000  # Number of steps to "tune" the samplers
burnInSteps <- 2000 
thinSteps <- 10
nChains <- 8
ncores <- detectCores()/2  # because of multi-threading, this is usually double the true number of physical cores, so we halve the number as Stan can only use physical cores
options(mc.cores=ncores)

if (!file.exists(paste0(datadir,'matchingdata.csv'))){
osfnode <- 'ey7rg'  
osfproject <- osf_retrieve_node(osfnode)
matchingfiles <- osf_ls_files(osfproject,n_max=300)
fid <- which(matchingfiles$name=='matchingdata.csv')
osf_download(matchingfiles[fid,],datadir,progress=TRUE)
}

matchingdata <- read.csv(paste0(datadir,'matchingdata.csv'))

matchingdatalist = list(correct = matchingdata$correct, allLratio = matchingdata$allLratio, allRratio = matchingdata$allRratio, allcontrast = matchingdata$allcontrast, allref = matchingdata$allref, s = matchingdata$Participant, Nsubj = max(matchingdata$Participant), Ntotal = nrow(matchingdata))

for (participant in 1:30){
         # if the tar file of this participant's data doesn't exist
         if (!file.exists(paste0(datadir,'P',100+participant,'summary.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name==paste0('P',100+participant,'summary.RData'))
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }
}

d <- dir(datadir,pattern='*summary.RData',full.names=TRUE)

allmasksP <- array(0,dim=c(length(d),6,5))
allmeansP <- allmasksP
allmasksE <- array(0,dim=c(length(d),64,6,5))
allmeansE <- allmasksE
allmeansE2 <- allmeansE
allmasksE2 <- allmasksE

  for (s in 1:length(d)){

    load(d[s])

  allmasksP[s,,] <- cleanmasksP
  allmeansP[s,,] <- cleanmeansP
  allmeansE[s,,,] <- cleanmeansE
  allmasksE[s,,,] <- cleanmasksE
  allmeansE2[s,,,] <- cleanmeansE2
  allmasksE2[s,,,] <- cleanmasksE2

  }
  electrodeindices <- match(targetelectrodes,electrodes)-2

  # THE DATA
  condlist <- c(1,2,3,5,6)
  contlist <- c(6,12,24,48,96)
  y <- NULL
  x <- NULL
  c <- NULL
  s <- NULL
  w <- NULL
  i <- 0
  for (subj in 1:dim(allmeansP)[1]){
    for (cond in 1:length(condlist)){
      for (cont in 1:length(contlist)){
        dp <- abs(allmeansP[subj,condlist[cond],cont])
        if (is.na(dp)==0){
          if (is.infinite(dp)==0){
          i <- i + 1
          x[i] <- contlist[cont]
          y[i] <- dp
          c[i] <- cond
          s[i] <- subj
          w[i] <- 1
          }
        }
      }
    }
  }

pupildatalist = list(x = x, y = y, s = s, w = w, c = c, Nsubj = max(s), Ntotal = length(y))

  y <- NULL
  x <- NULL
  c <- NULL
  s <- NULL
  w <- NULL
  i <- 0
  for (subj in 1:dim(allmeansE)[1]){
    for (cond in 1:length(condlist)){
      for (cont in 1:length(contlist)){
        dp <- abs(mean(allmeansE[subj,electrodeindices,condlist[cond],cont]))
        if (is.na(dp)==0){
          if (is.infinite(dp)==0){
            i <- i + 1
            x[i] <- contlist[cont]
            y[i] <- dp
            c[i] <- cond
            s[i] <- subj
            w[i] <- 1
          }
        }
      }
    }
  }

  EEG1Fdatalist = list(x = x, y = y, s = s, w = w, c = c, Nsubj = max(s), Ntotal = length(y))

  y <- NULL
  x <- NULL
  c <- NULL
  s <- NULL
  w <- NULL
  i <- 0
  for (subj in 1:dim(allmeansE)[1]){
    for (cond in 1:length(condlist)){
      for (cont in 1:length(contlist)){
        dp <- abs(mean(allmeansE2[subj,electrodeindices,condlist[cond],cont]))
        if (is.na(dp)==0){
          if (is.infinite(dp)==0){
            i <- i + 1
            x[i] <- contlist[cont]
            y[i] <- dp
            c[i] <- cond
            s[i] <- subj
            w[i] <- 1
          }
        }
      }
    }
  }

  EEG2Fdatalist = list(x = x, y = y, s = s, w = w, c = c, Nsubj = max(s), Ntotal = length(y))

# Translate Stan models to C++ and compile to DSO:
matchingmodel <- stan_model(model_code=matchingmodelstring)  
CRFmodel <- stan_model(model_code=CRFmodelstring)

# Get MC sample of posterior:
pupilsamples <- sampling(object=CRFmodel, 
                    data = pupildatalist, 
                    chains = nChains,
                    iter = (ceiling(numSavedSteps/nChains)*thinSteps+burnInSteps), 
                    warmup = burnInSteps, 
                    thin = thinSteps,
                    cores = getOption("mc.cores", 1L))

EEG1Fsamples <- sampling(object=CRFmodel, 
                    data = EEG1Fdatalist, 
                    chains = nChains,
                    iter = (ceiling(numSavedSteps/nChains)*thinSteps+burnInSteps), 
                    warmup = burnInSteps, 
                    thin = thinSteps,
                    cores = getOption("mc.cores", 1L))

EEG2Fsamples <- sampling(object=CRFmodel, 
                    data = EEG2Fdatalist, 
                    chains = nChains,
                    iter = (ceiling(numSavedSteps/nChains)*thinSteps+burnInSteps), 
                    warmup = burnInSteps, 
                    thin = thinSteps,
                    cores = getOption("mc.cores", 1L))

matchingsamples <- sampling(object=matchingmodel, 
                    data = matchingdatalist, 
                    chains = nChains,
                    iter = (ceiling(numSavedSteps/nChains)*thinSteps+burnInSteps), 
                    warmup = burnInSteps, 
                    thin = thinSteps,
                    cores = getOption("mc.cores", 1L))

save(file=paste0(datadir,'modeloutput.RData'),list=c('pupilsamples','EEG1Fsamples','EEG2Fsamples','matchingsamples'))

}

```

```{r include=FALSE, results='hide'}

# create figure containing modelling results

if (!file.exists(paste0(datadir,'modeloutput.RData'))){
osfnode <- '2qu6y'  
osfproject <- osf_retrieve_node(osfnode)
osffilesP <- osf_ls_files(osfproject,n_max=300)
fid <- which(osffilesP$name=='modeloutput.RData')
osf_download(osffilesP[fid,],datadir,progress=TRUE)
}

  load(paste0(datadir,'modeloutput.RData'))

  nchains <- dim(pupilsamples)[2]
  pupilweightpop <- NULL
  pupilZpop <- NULL
  pupilKpop <- NULL
  pupilRpop <- NULL
  for (n in 1:nchains){
  pupilweightpop <- c(pupilweightpop,pupilsamples@sim$samples[[n]]$wtmu)
  pupilZpop <- c(pupilZpop,pupilsamples@sim$samples[[n]]$Zmu)
  pupilKpop <- c(pupilKpop,pupilsamples@sim$samples[[n]]$kmu)
  pupilRpop <- c(pupilRpop,pupilsamples@sim$samples[[n]]$Rmaxmu)
  }
  pupilW <- median(pupilweightpop)
  pupilZ <- median(pupilZpop)
  pupilK <- median(pupilKpop)
  pupilRmax <- median(pupilRpop)

  nchains <- dim(EEG1Fsamples)[2]
  EEGweightpop <- NULL
  EEGZpop <- NULL
  EEGKpop <- NULL
  EEGRpop <- NULL
  for (n in 1:nchains){
  EEGweightpop <- c(EEGweightpop,EEG1Fsamples@sim$samples[[n]]$wtmu)
  EEGZpop <- c(EEGZpop,EEG1Fsamples@sim$samples[[n]]$Zmu)
  EEGKpop <- c(EEGKpop,EEG1Fsamples@sim$samples[[n]]$kmu)
  EEGRpop <- c(EEGRpop,EEG1Fsamples@sim$samples[[n]]$Rmaxmu)
  }
  EEGW <- median(EEGweightpop)
  EEGZ <- median(EEGZpop)
  EEGK <- median(EEGKpop)
  EEGRmax <- median(EEGRpop)

  nchains <- dim(EEG2Fsamples)[2]
  EEG2weightpop <- NULL
  EEG2Zpop <- NULL
  EEG2Kpop <- NULL
  EEG2Rpop <- NULL
  for (n in 1:nchains){
  EEG2weightpop <- c(EEG2weightpop,EEG2Fsamples@sim$samples[[n]]$wtmu)
  EEG2Zpop <- c(EEG2Zpop,EEG2Fsamples@sim$samples[[n]]$Zmu)
  EEG2Kpop <- c(EEG2Kpop,EEG2Fsamples@sim$samples[[n]]$kmu)
  EEG2Rpop <- c(EEG2Rpop,EEG2Fsamples@sim$samples[[n]]$Rmaxmu)
  }
  EEG2W <- median(EEG2weightpop)
  EEG2Z <- median(EEG2Zpop)
  EEG2K <- median(EEG2Kpop)
  EEG2Rmax <- median(EEG2Rpop)

  nchains <- dim(matchingsamples)[2]
  matchingweightpop <- NULL
  matchingZpop <- NULL
  matchingKpop <- NULL
  for (n in 1:nchains){
  matchingweightpop <- c(matchingweightpop,matchingsamples@sim$samples[[n]]$wmu)
  matchingZpop <- c(matchingZpop,matchingsamples@sim$samples[[n]]$Zmu)
  matchingKpop <- c(matchingKpop,matchingsamples@sim$samples[[n]]$kmu)
  }
  matchingW <- median(matchingweightpop)
  matchingZ <- median(matchingZpop)
  matchingK <- median(matchingKpop)

  if (processdata > 0){

postscript('posteriors.ps', horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

plotlims <- c(-48,12,0,0.6) 
ticklocsx <- seq(-48,12,12)    
ticklocsy <- seq(0,0.6,0.1)    
ticklabelsx <- c('1/256','1/64','1/16','1/4','1','4')        
ticklabelsy <- ticklocsy    # set labels for y ticks

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)

title(xlab="w", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Probability", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

a <- density(20*log10(pupilweightpop),from=-48,to=12)   
polygon(c(-48,a$x,12), c(0,a$y,0), col=colpal[3],border=NA)

a <- density(20*log10(EEGweightpop),from=-48,to=12)   
polygon(c(-48,a$x,12), c(0,a$y,0), col=colpal[4],border=NA)

a <- density(20*log10(EEG2weightpop),from=-48,to=12)   
polygon(c(-48,a$x,12), c(0,a$y,0), col=colpal[5],border=NA)

a <- density(20*log10(matchingweightpop),from=-48,to=12)   
polygon(c(-48,a$x,12), c(0,a$y,0), col=colpal[6],border=NA)

prior <- dnorm(10^((-48:12)/20),1,0.5)
prior <- 0.5*prior/max(prior)
lines(-48:12,prior)

lines(20*log10(c(mean(pupilW),mean(pupilW))),c(0,0.6),lwd=3,col=colpal[3])
lines(20*log10(c(mean(EEGW),mean(EEGW))),c(0,0.6),lwd=3,col=colpal[4])
lines(20*log10(c(mean(EEG2W),mean(EEG2W))),c(0,0.6),lwd=3,col=colpal[5])
lines(20*log10(c(mean(matchingW),mean(matchingW))),c(0,0.6),lwd=3,col=colpal[6])

legend(-48,0.6,c('Prior','Pupillometry','EEG 1F','EEG 2F','Matching'),col=c('black',colpal[3:6]),lwd=c(1,3,3,3,3),lty=1, box.lwd=2, bg='white')

text(9,0.58,'(i)',cex=1.25,adj=0.5)

dev.off()

finex <- 10^(seq(20*log10(6),20*log10(96),length=100)/20)

w <- pupilW
Z <- pupilZ
k <- pupilRmax
n <- pupilK
q <- 2
modelA <- matrix(0,nrow=6,ncol=length(finex))
modelB <- matrix(0,nrow=6,ncol=length(finex))
# mon condition
R <- 0
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[1,] <- k*(Lresp + Rresp) + n
modelB[1,] <- n

# bin condition
R <- finex
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[2,] <- k*(Lresp + Rresp) + n
modelB[2,] <- n

# dich condition
R <- 48
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[3,] <- k*(Lresp + Rresp) + n
modelB[3,] <- n

# mon X condition
R <- 0
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[4,] <- n
modelB[4,] <- k*(Lresp + Rresp) + n

# bin X condition
R <- finex
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[5,] <- k*(Lresp) + n 
modelB[5,] <- k*(Rresp) + n

# dich X condition
L <- finex
R <- 48
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[6,] <- k*(Lresp) + n 
modelB[6,] <- k*(Rresp) + n

contrastsdB <- 20*log10(c(6,12,24,48,96))

plotlims <- c(15,40,0,0.04)
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklocsy <- seq(0,0.04,0.01)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript(paste("CRF1p.ps",sep=''), horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)    
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Amplitude at 2Hz (mm)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

condlist <- c(1,2,3,6)
for (cond in 1:4){lines(20*log10(finex),modelA[condlist[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)}

legend(21,0.04,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),lwd=3,col=colpal[condlist],box.lwd=2)

text(17,0.038,'(e)',adj=0.5,cex=2.5)

dev.off()

w <- EEGW
Z <- EEGZ
k <- EEGRmax
n <- EEGK
q <- 2
modelA <- matrix(0,nrow=6,ncol=length(finex))
modelB <- matrix(0,nrow=6,ncol=length(finex))
# mon condition
R <- 0
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[1,] <- k*(Lresp + Rresp) + n
modelB[1,] <- n

# bin condition
R <- finex
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[2,] <- k*(Lresp + Rresp) + n
modelB[2,] <- n

# dich condition
R <- 48
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[3,] <- k*(Lresp + Rresp) + n
modelB[3,] <- n

# mon X condition
R <- 0
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[4,] <- n
modelB[4,] <- k*(Lresp + Rresp) + n

# bin X condition
R <- finex
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[5,] <- k*(Lresp) + n 
modelB[5,] <- k*(Rresp) + n

# dich X condition
L <- finex
R <- 48
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[6,] <- k*(Lresp) + n 
modelB[6,] <- k*(Rresp) + n

plotlims <- c(15,40,0,0.8)  
ticklocsy <- seq(0,0.8,0.2)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("CRF1e.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)    
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Amplitude at 2Hz (µV) ", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

condlist <- c(1,2,3,6)
for (cond in 1:4){lines(20*log10(finex),modelA[condlist[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)}

# legend(21,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),lwd=3,col=colpal[condlist],box.lwd=2)

text(17,0.75,'(f)',adj=0.5,cex=2.5)

dev.off()

w <- EEG2W
Z <- EEG2Z
k <- EEG2Rmax
n <- EEG2K
q <- 2
modelA <- matrix(0,nrow=6,ncol=length(finex))
modelB <- matrix(0,nrow=6,ncol=length(finex))
# mon condition
R <- 0
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[1,] <- k*(Lresp + Rresp) + n
modelB[1,] <- n

# bin condition
R <- finex
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[2,] <- k*(Lresp + Rresp) + n
modelB[2,] <- n

# dich condition
R <- 48
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[3,] <- k*(Lresp + Rresp) + n
modelB[3,] <- n

# mon X condition
R <- 0
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[4,] <- n
modelB[4,] <- k*(Lresp + Rresp) + n

# bin X condition
R <- finex
L <- finex
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[5,] <- k*(Lresp) + n 
modelB[5,] <- k*(Rresp) + n

# dich X condition
L <- finex
R <- 48
Lresp <- (L^q)/(Z + L + w*R)
Rresp <- (R^q)/(Z + R + w*L)
modelA[6,] <- k*(Lresp) + n 
modelB[6,] <- k*(Rresp) + n

plotlims <- c(15,40,0,0.8)  
ticklocsy <- seq(0,0.8,0.2)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("CRF2e.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)    
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)    
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude at 4Hz (µV) ", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

condlist <- c(1,2,3,6)
for (cond in 1:4){lines(20*log10(finex),modelA[condlist[cond],], col=colpal[condlist[cond]], lwd=3, cex=0.5)}

# legend(21,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),lwd=3,col=colpal[condlist],box.lwd=2)

text(17,0.75,'(g)',adj=0.5,cex=2.5)

dev.off()

# very finely sampled ratios to produce smooth curves
ratiosL <- seq(0,1,length=100)
ratiosR <- rep(1,100)

params <- c(2,matchingZ,matchingW)

binmodel <- function(L,R,p){
  # this is the two-stage model with a linear denominator and saturation constant
  stage1L <- (L^p[1])/(p[2] + L + p[3]*R)
  stage1R <- (R^p[1])/(p[2] + R + p[3]*L)
  resp <- stage1L + stage1R

  return(resp)}

getmatch <- function(L,R,ratios,p){
  standardresp <- binmodel(L,R,p)
  contrast <- 0
  matchresp <- 0
  while (matchresp<standardresp){
    contrast <- contrast + 0.1
    matchresp <- binmodel(contrast*ratios[1],contrast*ratios[2],p)
  }

  return(contrast)}

postscript("matchingmodel.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

plotlims <- c(0,125,0,125)  
ticklocs <- seq(0,125,25)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs)     # add the tick labels
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="Right eye contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Left eye contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

standardcontrast <- 48

matchlist <- NULL
for (n in 1:length(ratiosL)){
  matchlist[n] <- getmatch(standardcontrast,standardcontrast,c(ratiosL[n],ratiosR[n]),params)
}

tempx <- c(matchlist*ratiosL, matchlist[100:1])
tempy <- c(matchlist, matchlist[100:1]*ratiosL[100:1])

lines(tempx,tempy,lwd=3,col=colpal[2])
lines(c(0,standardcontrast),c(standardcontrast,standardcontrast),lty=2)
lines(c(standardcontrast,standardcontrast),c(0,standardcontrast),lty=2)
lines(c(0,2*standardcontrast),c(2*standardcontrast,0),lty=3)

standardcontrast <- 24

matchlist <- NULL
for (n in 1:length(ratiosL)){
  matchlist[n] <- getmatch(standardcontrast,standardcontrast,c(ratiosL[n],ratiosR[n]),params)
}

tempx <- c(matchlist*ratiosL, matchlist[100:1])
tempy <- c(matchlist, matchlist[100:1]*ratiosL[100:1])
lines(tempx,tempy,lwd=3,col=colpal[1])

lines(c(0,standardcontrast),c(standardcontrast,standardcontrast),lty=2)
lines(c(standardcontrast,standardcontrast),c(0,standardcontrast),lty=2)
lines(c(0,2*standardcontrast),c(2*standardcontrast,0),lty=3)

legend(62,95,c('48% match','24% match'),col=colpal[2:1],lwd=3,pt.cex=2,box.lwd=2)

legend(45,125,c('Winner-take-all','Linear summation'),lty=2:3,box.lwd=2)

text(10,117.5,'(h)',adj=0.5,cex=2.5)

dev.off()

contrastsdB <- 20*log10(c(6,12,24,48,96))

         if (!file.exists(paste0(datadir,'Averagedata.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='Averagedata.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }

load(file=paste0(datadir,'Averagedata.RData'))

plotlims <- c(15,40,0,0.03) 
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklocsy <- seq(0,0.03,0.01)    # locations of tick marks on y axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("pdata.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    # titles for axes
title(ylab="Amplitude at 2Hz (mm)", col.lab=rgb(0,0,0), line=1.8, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansP[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  lines(contrastsdB,abs(cleanmeansP[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves

legend(21,0.03,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),pch=c(21:23,25),pt.bg=colpal[c(1:3,6)],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.028,'(a)',adj=0.5,cex=2.5)

dev.off()

plotlims <- c(15,40,0,0.8)  
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklocsy <- seq(0,0.8,0.2)    # locations of tick marks on y axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("edata.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV) at 2Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansEav[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  lines(contrastsdB,abs(cleanmeansEav[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves

# legend(22,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),pch=c(21:23,25),pt.bg=colpal[c(1:3,6)],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(b)',adj=0.5,cex=2.5)

dev.off()

postscript("edata2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV) at 4Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansE2av[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  lines(contrastsdB,abs(cleanmeansE2av[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves

# legend(22,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),pch=c(21:23,25),pt.bg=colpal[c(1:3,6)],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.75,'(c)',adj=0.5,cex=2.5)

dev.off()

         if (!file.exists(paste0(datadir,'meanmatching.RData'))){
           # download it from OSF
           if (!exists('osffilesP')){
           osfnode <- '2qu6y'
           osfproject <- osf_retrieve_node(osfnode)
          osffilesP <- osf_ls_files(osfproject,n_max=300)}
          fid <- which(osffilesP$name=='meanmatching.RData')
          osf_download(osffilesP[fid,],datadir,progress=TRUE)
         }

load(paste0(datadir,'meanmatching.RData'))

postscript("matchingdata.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

plotlims <- c(0,125,0,125)  
ticklocs <- seq(0,125,25)   
ratiosL <- c(1, 1, 1, 1, 1, 0.75, 0.5, 0.25, 0)
ratiosR <- c(0, 0.25, 0.5, 0.75, 1, 1, 1, 1, 1)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocs, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocs, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocs, side = 1, at=ticklocs)     # add the tick labels
mtext(text = ticklocs, side = 2, at=ticklocs, line=0.2, las=1)
title(xlab="Right eye contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Left eye contrast (%)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

lines(c(0,24),c(24,24),lty=2)
lines(c(0,48),c(48,48),lty=2)
lines(c(24,24),c(0,24),lty=2)
lines(c(48,48),c(0,48),lty=2)
lines(c(0,96),c(96,0),lty=3)
lines(c(0,48),c(48,0),lty=3)

for (matchlev in 1:2){
lines(ratiosR*10^(meanthresh[matchlev,]/20),ratiosL*10^(meanthresh[matchlev,]/20),pch=16,col=colpal[matchlev],lwd=3)}

for (matchlev in 1:2){
for (cond in 1:9){
  points(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),pch=16,cex=2,col=colpal[matchlev])

  # if ((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])<42){
  # arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),ratiosR[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),ratiosL[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),length=0.015, angle=90, lwd=2)
  # }
  # 
  # # plot as an arrow if exceeding the axis limits
  # if ((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])>42){
  # arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),125,ratiosL[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),length=0.035, angle=45, lwd=2)
  # }

  # arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),ratiosR[cond]*10^((meanthresh[matchlev,cond]-SEthresh[matchlev,cond])/20),ratiosL[cond]*10^((meanthresh[matchlev,cond]-SEthresh[matchlev,cond])/20),length=0.015, angle=90, lwd=2)
}
}

text(10,117.5,'(d)',adj=0.5,cex=2.5)

legend(75,95,c('48% match','24% match'),pch=16,col=colpal[2:1],pt.cex=2,box.lwd=2)

legend(45.75,125,c('Winner-take-all','Linear summation'),lty=2:3,box.lwd=2)

dev.off()

  PostScriptTrace('CRF1p.ps')
  p1 <- readPicture('CRF1p.ps.xml')
  PostScriptTrace('CRF1e.ps')
  p2 <- readPicture('CRF1e.ps.xml')
  PostScriptTrace('CRF2e.ps')
  p3 <- readPicture('CRF2e.ps.xml')
  PostScriptTrace('matchingmodel.ps')
  p4 <- readPicture('matchingmodel.ps.xml')
  PostScriptTrace('posteriors.ps')
  p5 <- readPicture('posteriors.ps.xml')

  PostScriptTrace('pdata.ps')
  p6 <- readPicture('pdata.ps.xml')
  PostScriptTrace('edata.ps')
  p7 <- readPicture('edata.ps.xml')
  PostScriptTrace('edata2.ps')
  p8 <- readPicture('edata2.ps.xml')
  PostScriptTrace('matchingdata.ps')
  p9 <- readPicture('matchingdata.ps.xml')

    for (n in 1:length(p5@paths)){
    temp <- class(p5@paths[n]$path)[1]
    if (pmatch(temp,"PictureFill",nomatch=0)){
      if (sum(col2rgb(p5@paths[n]$path@rgb))<765){p5@paths[n]$path@rgb <- addalpha(p5@paths[n]$path@rgb,alpha=0.2)}}}

  pdf(paste0(figdir,"modelfigure.pdf"), bg="transparent", height = 10, width = 30)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  grid.picture(p1,x=1/12,y=0.25,width=1/6,height=1)
  grid.picture(p2,x=3/12,y=0.25,width=1/6,height=1)
  grid.picture(p3,x=5/12,y=0.25,width=1/6,height=1)
  grid.picture(p4,x=7/12,y=0.25,width=1/6,height=1)

  grid.picture(p5,x=5/6,y=0.5,width=1/3,height=1)

  grid.picture(p6,x=1/12,y=0.75,width=1/6,height=1)
  grid.picture(p7,x=3/12,y=0.75,width=1/6,height=1)
  grid.picture(p8,x=5/12,y=0.75,width=1/6,height=1)
  grid.picture(p9,x=7/12,y=0.75,width=1/6,height=1)

  dev.off()

  file.remove(c('CRF1e.ps','CRF2e.ps','CRF1p.ps','matchingmodel.ps','matchingdata.ps','pdata.ps','edata.ps','edata2.ps','posteriors.ps'))
  file.remove(c('CRF1e.ps.xml','CRF2e.ps.xml','CRF1p.ps.xml','matchingmodel.ps.xml','matchingdata.ps.xml','pdata.ps.xml','edata.ps.xml','edata2.ps.xml','posteriors.ps.xml'))

}

```

```{r modelfigure, fig.cap="Summary of computational modelling. Panels (a-d) show empirical data from key conditions, replotted from earlier figures for the pupillometry (a), first harmonic EEG responses (b), second harmonic EEG responses (c) and contrast matching (d) experiments. Panels (e-h) show model behaviour for the same conditions, generated using the median group-level parameter values.  Panel (i) shows the posterior probability distributions of the interocular suppression parameter for each of the four model fits. The pupillometry distribution (green) is centred about a substantially higher suppressive weight than for the other data types (note the logarithmic x-axis). The black curve shows the (scaled) prior distribution for the weight parameter.", fig.align="center", echo=FALSE}

knitr::include_graphics(paste0(figdir,'modelfigure.pdf'))

```

```{r paramtable, echo=FALSE}

exptnames <- c('Pupillometry','EEG 1F', 'EEG 2F', 'Matching')
zlist <- c(pupilZ,EEGZ,EEG2Z,matchingZ)
klist <- c(pupilK,EEGK,EEG2K,matchingK)
wlist <- c(pupilW,EEGW,EEG2W,matchingW)
Rlist <- c(round(c(pupilRmax,EEGRmax,EEG2Rmax),5),'-')
tabledata <- data.frame(exptnames,round(zlist,digits=2),round(klist,digits=2),round(wlist,digits=2),Rlist)

colnames(tabledata) <- c('Data set','Z','k','w','Rmax')
kable(tabledata, align='lcccc',caption = 'Summary of median parameter values.')

```

# Discussion

Using a novel paradigm that combines EEG and pupillometry, we found surprising results for the binocular integration of flickering light. In the visual cortex, the processing of spatially-uniform temporal luminance modulations seems to happen approximately linearly, with no evidence of interocular suppression. In the subcortical pathway, signal combination is more non-linear, with evidence of interocular suppression. This pattern of results was confirmed by computational modelling, which showed a much greater suppressive weight for the pupillometry data compared to the EEG data. Additionally, we found that perception of flickering light is consistent with a near-linear summation process, consistent with the cortical (EEG) responses.

The results of our main experiment were unexpected for both the pupillometry and the EEG measures. Previous studies investigating binocular combination of spatial patterns (i.e. sine wave grating stimuli) are universally consistent with strong interocular suppression and weak binocular facilitation at high contrasts [@Baker2017; @Moradi2009; @Meese2006]. Our second experiment ruled out the possibility that these differences were due to the lower temporal frequency (2Hz) used here. However, there is evidence of more extensive binocular facilitation for a range of other stimuli. For example, Quaia et al.@Quaia2018 observed a strong binocular facilitation (or 'supersummation') in the reflexive eye movement response to rapidly moving stimuli, and Spitschan and Cajochen@Spitschan2019 report a similar result in archival data on melatonin suppression due to light exposure. In the auditory system, interaural suppression of amplitude modulation also appears to be weak when measured using a similar steady-state paradigm [@Baker2020]. Finally, psychophysical matching experiments using static stimuli also show near-linear behaviour for luminance increments [@Levelt1965; @Anstis1998; @Baker2012], though not for luminance decrements [@Anstis1998]. Overall, this suggests that strong interocular normalization may be specific to spatial pattern vision, and not a general feature of binocular signal combination.

Given the above, where does this leave our understanding of the overarching purpose of signal combination? We previously@Baker2017 pointed out that a model with strong suppression between channels that are subsequently summed is equivalent to a Kalman filter, which is the optimal method for combining two noisy inputs (see also [@Ernst2002]). This account has intuitive appeal, and is consistent with other models that propose binocular combination as a means of redundancy reduction [@Li1994; @May2022]. One possibility is that optimal combination is useful for visual perception --- a critical system for interacting with the local environment --- and is therefore worth devoting the additional resource of inhibitory wiring between ocular channels. However the other examples given in the previous paragraph are primarily physiological responses (pupil size, eye movements, hormone release) that may benefit less from an increased signal-to-noise ratio, or are otherwise phylogenetically older than binocular pattern vision. Conceptualised another way, the brain can repurpose a generic architecture for different situational demands by adjusting parameter values (here the weight of interocular suppression) to achieve different outcomes. Our future work in this area intends to compare binocular combination for specific photoreceptor pathways, including different cone classes, and intrinsically photoreceptive retinal ganglion cells.

Pupil size determines the total amount of light falling on the retina. It is therefore the case that fluctuations in pupil diameter have a downstream effect on the signals reaching cortex. We did not incorporate such interactions into our computational model, though in principle this might be worthwhile. However we anticipate than any such effects would be small, since pupil modulations at 2Hz are in the order of 2% of overall diameter [@Spitschan2014]. It is also the case that cortical activity can modulate pupil diameter, usually through arousal and attention mechanisms [@Bradley2008]. We think it unlikely that these temporally coarse processes would have a differential effect on e.g. monocular and binocular stimulation conditions in our experiment, and any fluctuations during an experimental session (perhaps owing to fatigue) will be equivalent for our comparisons of interest. Therefore we make the simplifying assumption that the two pathways are effectively distinct, but intend to investigate this directly in future neuroimaging work.

Classic studies investigating the neurophysiological architecture of V1 reported that cells in cytochrome-oxidase 'blobs' [@Horton1981; @Livingstone1984] are biased towards low spatial frequencies [@Tootell1988; @Edwards1995], and relatively insensitive to stimulus orientation [@Horton1981; @Livingstone1984] (though see @Economides2011). As the blob regions are embedded within ocular dominance columns [@Horton1981], they are also largely monocular [@Livingstone1984; @Tychsen2004]. More recent work has reported psychophysical evidence for unoriented chromatic [@Gheiratmand2013] and achromatic [@Meese2011] mechanisms, that also appear to be monocular. Our use of luminance flicker might preferentially stimulate these mechanisms, perhaps explaining why our EEG data show little evidence of binocular interactions. Indeed, our EEG results could potentially be explained by a model involving entirely non-interacting monocular channels, with the binocular facilitation effects we find (e.g. Fig. \@ref(fig:BSratios) & \@ref(fig:TFdata)) owing to additivity of the electrophysiological response across independent monocular cells, rather than requiring binocular neurons. However our matching data (Fig. \@ref(fig:matchingdata)), as well as everyday perceptual experience, indicates that luminance signals must be combined physiologically at some stage, and that this process still involves weak or absent interocular suppression.

# Conclusions

We have demonstrated that binocular combination of flickering light differs between cortical and subcortical pathways. Flicker was also associated with substantially weaker interocular suppression, and stronger binocular facilitation, compared to combination of spatial luminance modulations in visual cortex. Our computational framework for understanding signal combination permits direct comparisons between disparate experimental paradigms and data types. We anticipate that this will help elucidate the constraints the brain faces when combining different types of signals to govern perception, action and biological function.

# Methods

## Participants

Thirty (20 F, 10 M), twelve (7 F, 5 M) and ten (3 F, 7 M) adult participants, whose ages ranged from 18 to 45, were recruited for Experiments 1, 2 and 3 respectively. All participants had normal or corrected-to-normal binocular vision, and gave written informed consent in accordance with the Declaration of Helsinki. Our procedures were approved by the Ethics Committee of the Department of Psychology at the University of York (identification number 792).

## Apparatus & stimuli

The stimuli were two monocularly presented discs of achromatic flickering light with a diameter of 3.74 degrees, displayed against a black background. The same stimuli were used for all three experiments. Four dark red lines were added around each monocular disc to help with their fusion into a single binocular disc (see insert in Fig. \@ref(fig:pupildata)b for an example of the fused stimulus). The discs were viewed through a four-mirror stereoscope, which used front silvered mirrors to avoid internal reflections, and meant that participants perceived a single fused disc. The use of a stereoscope allowed us to modulate the stimuli in three different ocular configurations: monocular, binocular, and dichoptic. Note that during monocular presentation of flicker, the unstimulated eye still saw the static (non-flickering) disc of mean luminance.

All stimuli had a mean luminance of 42 cd/m$^2$, and were presented on an Iiyama VisionMaster™ Pro 510 display (800 x 600 pixels, 60 Hz refresh rate), which was gamma corrected using a Minolta LS-110 photometer (Minolta Camera Co. Ltd., Japan). All stimuli were presented using Psychopy [@Peirce2019] (v3.0.7 for Experiments 1 and 2; v2022.1.1 for Experiment 3).

EEG data were collected for Experiments 1 and 2 using a 64 electrode ANT WaveGuard cap and the signals were recorded at 1 kHz using the ASA software (ANT Neuro, Netherlands). Pupillometry data were collected for Experiment 1 using a binocular Pupil Core eye-tracker (Pupil Labs GmbH, Berlin, Germany @Kassner2014) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

## Procedure

Before each experiment, participants calibrated the stereoscope by adjusting the angle of the mirrors. This was done so that they would perceive the two discs as one fused disc when looking at the screen through the stereoscope.

### Experiment 1: simultaneous EEG and pupillometry

The experiment was conducted in a windowless room, in which the only light source was the monitor. The participants sat at 99 cm from the monitor and the total optical viewing distance (through the stereoscope) was 107 cm. The experiment was carried out in a single session lasting 45 minutes in total, divided in three blocks of 15 minutes each. In each block, there were 60 trials lasting 15 seconds each (12s of stimulus presentation, with an interstimulus interval of 3s). The participants were given no task other than to look at the fixation cross in the middle of the disc while trying to minimise their blinking during the presentation period.

We included six distinct ocular conditions, each at five temporal contrast levels (combined factorially) relative to the mean luminance: 6, 12, 24, 48 and 96\%. Contrast was defined as temporal Michelson contrast; the difference between maximum and minimum luminances, scaled by the mean and expressed as a percentage. In the first three conditions, the discs flickered at 2 Hz, in either a monocular, binocular, or dichoptic arrangement. In the dichoptic condition the non-target eye saw a fixed contrast of 48\%. In the remaining three conditions (termed the cross-frequency conditions) one eye's disc flickered at 2Hz, and the other eye's disc flickered at 1.6Hz. We also tested monocular (one eye sees 1.6Hz flicker, the other sees mean luminance), binocular (one eye sees each frequency at the target contrast) and dichoptic (target stimulus flickering at 2Hz, mask contrast of 48\% at 1.6Hz in the other eye) arrangements. The rationale for flickering both eyes at 2Hz is that we can measure summation behaviour between the eyes in the pupil and EEG response at 2Hz. The rationale for flickering the eyes at different frequencies is that this permits measurement of suppression between the eyes (i.e. the reduction in the 2Hz response when a 1.6Hz mask component is added to the other eye). We counterbalanced presentation of the target stimulus across the left and right eyes.

### Experiment 2: EEG responses across temporal frequency

This experiment used the same equipment set up as Experiment 1, except that the eye tracker was not used. Unlike the first experiment, only one contrast level was used (96\%) and the discs were set to flicker at five different frequencies (2, 4, 8, 16 and 30 Hz). Only two ocular configurations, monocular and binocular, were included, with the latter having both discs flickering at the same frequency. The experiment was carried out in one session lasting 25 minutes in total, divided into five blocks of 5 minutes each. In each block, there were 20 trials in total with the same timing as for Experiment 1.

### Experiment 3: temporal contrast matching

The experiment was conducted in a darkened room with a blacked-out window. The display equipment (monitor and stereoscope) were the same as for the two previous experiments, but no EEG or pupillometry data were collected. A two-interval contrast matching procedure was used to collect data. In one interval, participants were presented with a standard fused disc that flickered at a set contrast level (either 24 or 48\%), which was selected by the experimenter at the beginning of each block. In the other interval, a target disc flickering at varying contrast levels was displayed. The contrast level of the target was controlled by a 1-up, 1-down staircase moving in logarithmic (dB) steps of contrast. The ratio of flicker amplitudes in the left and right eyes was varied across blocks and was set to be 0, 0.25, 0.5, 0.75 or 1 (9 distinct conditions). The standard and target discs were displayed for 1 second each, with an interstimulus interval of 0.5 seconds. After both discs had appeared on screen, the participants had to indicate which interval they perceived as having the more intense flicker. The intervals were randomly ordered, and all discs flickered at a frequency of 2 Hz (two cycles in sine phase).

Due to its long duration (approximately 3 hours in total), the participants completed the experiment across multiple sessions initiated at their own convenience. The experiment was divided into 54 blocks (3 repetitions $\times$ 2 standard contrasts $\times$ 9 target ratios), which lasted on average 3 minutes each, depending on the response speed of the participant. In each block, there were a total of 50 trials. No auditory feedback was given for this subjective task.

## Data analysis

EEG data were converted from the ANT-EEProbe format to a compressed csv text file using a custom Matlab script and components of the EEGlab toolbox [@Delorme2004]. The data for each participant were then loaded into R for analysis, where a ten-second waveform for each trial at each electrode was extracted (omitting the first two seconds). The Fourier transform of each waveform was calculated, and the complex spectrum stored in a matrix. All repetitions of each condition were then averaged for each electrode. They were then averaged across four occipital electrodes (\emph{POz}, \emph{Oz}, \emph{O1}, \emph{O2}), to obtain individual results. Finally, these were averaged across participants to obtain the group results. All averaging was performed in the complex domain and therefore retained the phase information (i.e. coherent averaging), and at each stage we excluded data points with a Mahalanobis distance exceeding $D$ = `r SDthresh` from the complex-valued mean [@Baker2021]. For statistical comparisons of complex-valued data, we use the $ANOVA^2_{circ}$ statistic described previously @Baker2021. This is a multivariate extension of ANOVA that assumes equal variance of the real and imaginary Fourier components, or equivalently, an extension of the $T^2_{circ}$ statistic @Victor1991 that can compare more than two conditions.

A similar analysis pipeline was adopted for the pupillometry data. The data were converted from mp4 videos to a csv text file using the Pupil Player software [@Kassner2014], which estimated pupil diameter for each eye on each frame using a 3D model of the eyeball. The individual data were then loaded into R for analysis, where again a ten-second waveform for each trial in each eye was extracted (excluding the first two seconds after stimulus onset). We interpolated across any dropped or missing frames to ensure regular and continuous sampling over time. The Fourier transform was calculated for each waveform, and all repetitions of each condition were pooled across eye and then averaged. Finally, data were averaged across all participants to obtain the group results. Again, we used coherent averaging, and excluded outlying data points in the same way as for the EEG data. Note that previous pupillometry studies using luminance flicker have tended to fit a single sine-wave at the fundamental frequency, rather than using Fourier analysis @Spitschan2014. The Fourier approach is more robust to noise at other frequencies and has been used in some previous studies [@Barrionuevo2014; @Barrionuevo2016]. Additionally, it makes the pupillometry analysis consistent with standard practice in steady-state EEG analysis [@Figueira2022].

To analyse the matching data, we pooled the trial responses across all repetitions of a given condition for each participant. We then fitted a cumulative normal psychometric function to estimate the point of subjective equality at the 50\% level. Thresholds were averaged across participants in logarithmic (dB) units.

For all experiments, we used a bootstrapping procedure with `r nbootstraps` iterations to estimate standard errors across participants. All analysis and figure construction was conducted using a single R-script, available online, making this study fully computationally reproducible.

## Computational model and parameter estimation

To describe our data, we chose a model of binocular contrast gain control with the same general form as the first stage of the model proposed by Meese et al. [@Meese2006]. The second gain control stage was omitted, consistent with previous work [@Baker2017], to simplify the model and reduce the number of free parameters. The response of the left eye's channel is given by:

\begin{equation}
\label{eq:respL}
Resp_L = \frac{L^2}{Z + L + wR},
\end{equation}

with an equivalent expression for the right eye:

\begin{equation}
\label{eq:respR}
Resp_R = \frac{R^2}{Z + R + wL}.
\end{equation}

In both equations, _L_ and _R_ are the contrast signals from the left and right eyes, _Z_ is a saturation constant that shifts the contrast-response function laterally, and _w_ is the weight of suppression from the other eye.

The responses from the two eyes are then summed binocularly:

\begin{equation}
\label{eq:respB}
Resp_B = R_{max}(Resp_L + Resp_R) + n,
\end{equation}

where _n_ is a noise parameter, and $R_{max}$ scales the overall response amplitude. The $R_{max}$ parameter was omitted when modelling the contrast matching data, as it has no effect in this paradigm.

Despite being derived from the model proposed by Meese et al. [@Meese2006], the simplifications applied to this architecture make it very similar to other models [@Ding2006; @Legge1984; @Schrodinger1926; @Doesschate1967]. In particular we fixed the numerator exponent at 2 in our model, because otherwise this value tends to trade off with the weight of interocular suppression [@Baker2012; @Kingdom2015]. Our key parameter of interest is the weight of interocular suppression. Large values around $w$ = 1 result in a very small or nonexistent binocular advantage at suprathreshold contrasts, consistent with previous work using grating stimuli [@Baker2017]. Low values around $w$ = 0 produce substantial, near-linear binocular facilitation [@Baker2020].

We implemented the model within a Bayesian framework using the Stan software [@Carpenter2017]. This allowed us to estimate group-level posterior parameter distributions for the weight of interocular suppression, $w$, and the other free model parameters $R_{max}$, $Z$ and $n$. The prior distributions for all parameters were Gaussian, with means and standard deviations of 1 and 0.5 for $w$ and $R_{max}$, and 5 and 2 for $Z$ and $n$, with these values chosen based on previous literature [@Meese2006; @Baker2012]. We sampled from a Student's t-distribution for the amplitudes in the pupillometry and EEG experiments, and from a Bernoulli distribution for the single trial matching data. The models were fit using the individual data across all participants, independently for each data set. Because we used coherent averaging across participants for our steady-state data, the group average amplitudes are shifted vertically relative to the model predictions, which are based on hierarchical fits to the individual participant amplitudes (put another way, the model does not implement coherent averaging). However it is clear that the character of the model in all cases gives a good representation of the data, as can be seen in Fig. \@ref(fig:modelfigure). We took posterior samples at over a million steps for each data set, using a computer cluster, and retained 10\% of samples for plotting.

## Preregistration

We initially preregistered our main hypotheses and analysis intentions for the first experiment. We then conducted a pilot study with N=12 participants, before making some minor changes to the stimulus (we added dim red lines to aid binocular fusion). We then ran the main experiment, followed by two additional experiments that were not preregistered. The preregistration document is available at the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

# Data availability

Raw data files are available at the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

# Code availability

Experimental and analysis code is available at the project repository: https://doi.org/10.17605/OSF.IO/TBEMA.

# Funding

Supported by Biotechnology and Biological Sciences Research Council grant BB/V007580/1 awarded to DHB and ARW, and Wellcome Trust grant 213616/Z/18/Z to AB.

# Author information

## Authors and Affiliations

**Department of Psychology, University of York, York, UK**

Federico G. Segala, Aurelio Bruno, Myat T. Aung, Alex R. Wade &  Daniel H. Baker

## Contributions

**F.G.S.**: Methodology, software, formal analysis, investigation, data curation, writing - original draft, writing - review & editing, visualization. **A.B.**: Conceptualization, writing - review & editing, supervision, project administration, funding acquisition. **M.A.**: Software, resources, writing - review & editing. **A.R.W.**: Conceptualization, methodology, resources, writing - review & editing, supervision, project administration, funding acquisition. **D.H.B.**: Conceptualization, methodology, software, formal analysis, investigation, resources, data curation, writing - original draft, writing - review & editing, visualization, supervision, project administration, funding acquisition.

## Corresponding Author

Correspondance to Federico G. Segala.

# Ethics declarations

## Competing interests

The authors declare no competing interests.

# References